---
title: "Projet_Women_FIFA_WC23_Analysis"
institute : "INSA Toulouse"
date: "`r Sys.Date()`"
always_allow_html: true
output: 
  pdf_document :
    toc : TRUE
    toc_depth : 3
    number_section : TRUE
    fig_caption: yes
header-includes:
   - \usepackage{dsfont}
   - \usepackage{color}
   - \newcommand{\1}{\mathds{1}}
---

```{r, include=FALSE}
#Eval=False : on compile pas le chunk
#Include=False : on affiche rien
#Echo=False : on affiche que la sortie
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(StatsBombR)
library(extrafont)
library(MASS)
library(dplyr)
library(plotly)
library(stringr)
library(ggplot2)
library(writexl)
library(readxl)
library(gridExtra)




```

\newpage

# Introduction

  Blalblablabla
  
  
```{r, include=FALSE}
# World Cup 2023 (id 72)
WC2023 <- FreeCompetitions() %>%
filter(competition_id==72 & season_name=="2023")

# Games availables in WC2023
Matches <- FreeMatches(WC2023)
```

```{r, include=FALSE,eval=F}

# Events for the games of the WC2023
WC2023_dataframe <- free_allevents(MatchesDF = Matches, Parallel = T)
WC2023_dataframe = allclean(WC2023_dataframe)

```


```{r, include=F}

# Lire le fichier CSV dans R
#write.csv(WC2023_dataframe, "WC2023_dataframe.csv", row.names = FALSE)

WC2023_dataframe <- read.csv("WC2023_dataframe.csv")

```



```{r,, include=FALSE}
WC2023_dataframe <- WC2023_dataframe %>%
  mutate(team.name = str_remove(team.name, " Women's"))

WC2023_dataframe <- WC2023_dataframe %>%
  mutate(possession_team.name = str_remove(possession_team.name, " Women's"))
```

# Descriptive data analysis 

 We begin by interpreting the elements of the data set.\
It is made up of multiple observations with different variables.\

## Analysis of successful shots according to country
   First, we look at the number of goals and shots in all matches for each team.\
Figure \ref{fig:fig1} shows a visualization of these results.\


```{r, include=FALSE}
#Number of goals and shots in all games for each team
shots_goals = WC2023_dataframe %>%
group_by(team.name) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 
```

```{r fig1 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig1}Diagram of the number of goals and shots in all matches for each team",fig.height=4.5}
#Let's make a graph
ggplot(shots_goals, aes(y = reorder(as.factor(team.name), goals))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of goals and shots in all games for each team",
       y = "Team",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green"))
```

It would be interesting to make this graph on the average number of goals and shots, as some teams have more games than others, distorting the results a little.

Figure \ref{fig:fig2} shows a visualization of the percentage of shots leading to a goal.\

```{r fig2 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig2}Diagram of the percentage of shots leading to a goal",fig.height=4,fig.align='center'}

ggplot(shots_goals, aes(y = reorder(as.factor(team.name), goals/shots*100))) +
  geom_bar(aes(x = goals/shots*100), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Percentage of shots leading to a goal",
       y = "Team",
       x = "Percentage")
```

\vspace{3pt}

We now would like to focus on France team.\
Figure \ref{fig:fig3} shows the results.\

```{r,include=F}
#Number of goals and shots in each match for both teams
shots_goals_all_matches = WC2023_dataframe %>%
group_by(match_id) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 

```

```{r,include=F}
#We display ids of all France matches
Id_France = WC2023_dataframe%>% filter(team.name=="France")
match_ids_france=unique(Id_France$match_id)
print(match_ids_france)
```


```{r,include=F}

#For each French match, the total number of shots and goals for the 2 teams (France+adversary) is displayed.


france_goals <- shots_goals_all_matches %>%
  filter(match_id %in% match_ids_france)

print(france_goals)
```

```{r,include=F}

ggplot(france_goals, aes(y = as.factor(match_id))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of shots and goals for each French match",
       y = "Match_id",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green"))
  
```



```{r,include=F}
ggplot(france_goals, aes(y = reorder(as.factor(match_id), goals/shots*100))) +
  geom_bar(aes(x = goals/shots*100), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Percentage of shots on goal in France matches",
       y = "Match",
       x = "Percentage")

#Il faut ordonner par date pour avoir l'évolution de la stratégie au cours de la compétition


```

```{r,include=F}

#We want to know which goals come from France
shots_goals_France_all_matches = WC2023_dataframe %>%
filter(team.name=="France") %>%
group_by(match_id) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 

print (shots_goals_France_all_matches)

```



```{r fig3 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig3}Diagram of the number of shots and goals for each French and percentage",fig.height=2,fig.align='center'}

g1=ggplot(shots_goals_France_all_matches, aes(y = as.factor(match_id))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Numbers for French matches",
       y = "Match_id",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green")) 


g2=ggplot(shots_goals_France_all_matches, aes(y = as.factor(match_id))) +
  geom_bar(aes(x = goals/shots*100), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Percentage in France matches",
       y = "Match",
       x = "Percentage") 

#Order from bottom to top: 1st match to last
grid.arrange(g1,g2,ncol=2)
```
\clearpage
 
## Analysis of successful shots according to different variables  

We first look at the type of shots.\

```{r,include=F}
#Number of goals and shots in all matches by type of shot
shots_goals_formation = WC2023_dataframe %>%
group_by(shot.type.name) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 

shots_goals_formation=shots_goals_formation[-c(5),]
```

```{r fig4 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig4}Diagram of the number of shots and goals for each type of shot",fig.height=2,fig.align='center'}
ggplot(shots_goals_formation, aes(y = reorder(as.factor(shot.type.name), goals))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of shots and goals for each type of shot",
       y = "type",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green"))
```


  We know would like to see if the technique of shot is significant.\

```{r,include=F}
#Number of shots and goals for each technique of shot
shots_goals_technique = WC2023_dataframe %>%
group_by(shot.technique.name) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 

#On enlève la ligne NA

shots_goals_technique=shots_goals_technique[-c(8),]
```

```{r fig5 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig5}Diagram of the number of shots and goals for each technique of shot",fig.height=2,fig.pos="h",fig.align='center'}

ggplot(shots_goals_technique, aes(y = reorder(as.factor(shot.technique.name), goals))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of shots and goals for each technique of shot",
       y = "Technique of shot",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green"))
```

Finally, is the variable body_zone_used relevant ?

```{r,include=F}
#Number of goals and shots over all matches according to body zone used
shots_goals_part = WC2023_dataframe %>%
group_by(shot.body_part.name) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 
shots_goals_part=shots_goals_part[-c(5),]

```

```{r fig6 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig6}Diagram of the number of goals and shots according to body zone used",fig.height=2,fig.pos="h"}

ggplot(shots_goals_part, aes(y = reorder(as.factor(shot.body_part.name), goals))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of goals and shots according to body zone used",
       y = "Body zone",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green")) 
```
\clearpage

# Models analysis

We wanted to create our own xG model. To do that we developed different models, finding the most relevant variables to predict goals.\

```{r,include=F}
#We create a df with just the shots

shots = WC2023_dataframe %>%
 filter(type.name=="Shot")
```

We run a logistic regression model: we want the output to be 0 or 1 depending on whether the shot turns into a goal.\

The first model keeps the variables studied previously : body part, technique, type of shot. \


```{r,include=F}

df_model_1 <- dplyr::select(shots,shot.outcome.name,shot.body_part.name,shot.technique.name,shot.type.name)
df_model_1$shot.outcome.name <- ifelse(df_model_1$shot.outcome.name == "Goal", 1, 0)


```

```{r, include = F}
#Testing regression without interaction

glm.model_1<-glm(shot.outcome.name~ . ,data=df_model_1,family=binomial(link="logit"))
summary(glm.model_1)
```

```{r,include=F}
#We see a skewed shot.type.namePenalty because the corner modality is taken as reference and, given that there were no shots or goals during the competition, the model considers 100% success for the corner modality.


#We change the reference modality from “corner” to “penalty”.
df_model_1_modif=df_model_1
df_model_1_modif$shot.type.name <- relevel(as_factor(df_model_1_modif$shot.type.name), ref = "Penalty")

glm.model_1_modif <- glm(shot.outcome.name ~ . , data = df_model_1_modif, family = binomial(link = "logit"))
summary(glm.model_1_modif)

```

```{r,include=F}
#We can see that Open Play and Free Kick are less likely to result in a goal than a penalty kick. The corner remains distorted
```


$R^2$ for the model without interaction is :\

```{r,echo=F}
#R2 calculation:
print(1-(glm.model_1$deviance/glm.model_1$null.deviance))

```

```{r,include=F}
#Testing regression with interaction
glm.model_1_interac<-glm(shot.outcome.name~ .^2 ,data=df_model_1,family=binomial(link="logit"))
summary(glm.model_1_interac)
#A lot of NA because some interactions are impossible (e.g. volleyball restart on a penalty kick).

```
$R^2$ for the model with interaction is :\

```{r,echo=F}
#R2 calculation:
print(1-(glm.model_1_interac$deviance/glm.model_1_interac$null.deviance))
```

We make a model with the given expected goal as variable.\


```{r,include=F}
df_model_2 <- dplyr::select(shots,shot.statsbomb_xg,shot.outcome.name)
df_model_2$shot.outcome.name <- ifelse(df_model_2$shot.outcome.name == "Goal", 1, 0)

glm.model_2<-glm(shot.outcome.name~ . ,data=df_model_2,family=binomial(link="logit"))
summary(glm.model_2)

```
We would therefore like to find a model with an $R^2$ value close to this model, i.e. an $R^2$ close to :
```{r,echo=F}
print(1-(glm.model_2$deviance/glm.model_2$null.deviance))

```

We do the same logistical model but with the position added : location.x and location.y

```{r,include=F}

df_model_3 <- dplyr::select(shots,shot.outcome.name,shot.body_part.name,shot.technique.name,shot.type.name,location.x,location.y)
df_model_3$shot.outcome.name <- ifelse(df_model_3$shot.outcome.name == "Goal", 1, 0)

```

```{r,include=F}
#Model without interaction
glm.model_3<-glm(shot.outcome.name~ . ,data=df_model_3,family=binomial(link="logit"))
summary(glm.model_3)
```

We test a regression without interaction, and obtain an $R^2$ of :

```{r,echo=F}
print(1-(glm.model_3$deviance/glm.model_3$null.deviance))

```

With interaction, and get an $R^2$ of :
```{r,include=F}
#Model with interaction

glm.model_3_interac<-glm(shot.outcome.name~ .^2 ,data=df_model_3,family=binomial(link="logit"))
summary(glm.model_3_interac)
```

```{r,echo=F}
print(1-(glm.model_3_interac$deviance/glm.model_3_interac$null.deviance))
```
With interactions, we targeted the main variables to obtain a good model and an $R^2$ as close to 1 as possible.


We run several tests to see which variables are significant in the model. 

```{r,echo=F}

model_simp<-glm(shot.outcome.name~ (location.x +location.y + shot.body_part.name+shot.type.name)^2 ,data=df_model_3,family=binomial(link="logit"))
anova(model_simp,glm.model_3_interac,test="Chisq")

```

We see that we can remove the technique because p_value>0.05 so we can accept the sub-model with a 95% level.

```{r,include=F}
summary(model_simp)
```

For this model we obtain an $R^2$ of :

```{r,echo=F}
print(1-(model_simp$deviance/model_simp$null.deviance))
```

The $R^2$ is no greater than for model 3 with interactions: this is normal because the $R^2$ favors models with many variables.
We should look at other variables such as AIC, which is minimum for model 3 without interactions.

```{r,include=F}
#We'll compare the fitted values (probability of being a goal), with the expected goals.
glm.model_3$fitted.values
#we have numbers between 0 and 1 so this is good.

```