---
title: "Projet_Women_FIFA_WC23_Analysis"
institute : "INSA Toulouse"
date: "`r Sys.Date()`"
always_allow_html: true
output: 
  pdf_document :
    toc : TRUE
    toc_depth : 3
    number_section : TRUE
    fig_caption: yes
header-includes:
   - \usepackage{dsfont}
   - \usepackage{color}
   - \newcommand{\1}{\mathds{1}}
---

```{r, include=FALSE}
#Eval=False : on compile pas le chunk
#Include=False : on affiche rien
#Echo=False : on affiche que la sortie
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(cowplot)
library(patchwork)
library(tidyverse)
library(StatsBombR)
library(extrafont)
library(MASS)
library(dplyr)
library(plotly)
library(stringr)
library(ggplot2)
library(writexl)
library(readxl)
library(gridExtra)
library(matrixStats)
library(ggplot2)
library(reshape2)
library(viridis)
```

\newpage

# Introduction

  In today's world, sports are at the center of global culture. To continue to excel, players and teams must find solutions, both physical and tactical. Therefore, statistics will play a crucial role in optimizing performance. Previous studies have shed light on various aspects of football analytics. Collet studied the impact of possession in 2013. More recently Liu analyzed the environmental impact in 2021. However, the realm of soccer remains relatively untapped in terms of data exploration. Understanding the dynamics of offensive and defensive play is pivotal for teams aiming to excel in competitions. 

The research gap lies in the need for a comprehensive analysis of football performance using advanced statistical methods, with a focus on data from platforms like StatsBomb. The impact of certain specific aspects of football analytics, such as shot analysis or passing patterns, remains unclear, and a comprehensive understanding of player and team performance is still lacking.

We aimed to address this gap by conducting a detailed analysis of football performance using StatsBomb data. We seeked to identify key performance indicators, assess their impact on match outcomes, and uncover underlying trends and patterns in player and team performance. This report outlines the methodology used for data collection and analysis, presents the findings from the study, and discusses their implications for the future of football analytics.

This report is divided into three parts. In the first section, we conduct an exploratory data analysis to identify certain trends, notably by analyzing shots and goals for each team. Then we seek an optimal statistical model to determine which parameters have the greatest impact on player performance. The last section contains the results of our analysis, including insights into player and team performance derived from StatsBomb data, with graphs examining successful shots and passes. 

  
  
```{r, include=FALSE}
# World Cup 2023 (id 72)
WC2023 <- FreeCompetitions() %>%
filter(competition_id==72 & season_name=="2023")

# Games availables in WC2023
Matches <- FreeMatches(WC2023)
```



```{r, include=FALSE}
AllMatches <- FreeMatches(WC2023)

AllMatches <- AllMatches  %>%  mutate(away_team.away_team_name = gsub(" Women's","",away_team.away_team_name),
                                      home_team.home_team_name = gsub(" Women's","",home_team.home_team_name),
                                      .keep="unused")


AllMatches <- AllMatches  %>%  mutate(away_team.away_team_name = gsub(" ","",away_team.away_team_name),
                                      home_team.home_team_name = gsub(" ","",home_team.home_team_name),
                                      .keep="unused")

Events <- free_allevents(AllMatches)
Shots_All <- Events %>% filter(type.name=="Shot")

Shots_All <- Shots_All  %>%  mutate(team = gsub(" ","",team.name),
                                      .keep="unused")
#view(Shots_All)
```

```{r, include=FALSE}
Shots <- subset(Shots_All, select = c(shot.outcome.name,team, player.name))

Shots <- Shots  %>%  mutate(player = player.name,
                                      outcome = case_when(shot.outcome.name!="Goal" ~ 0,
                                                         shot.outcome.name=="Goal" ~ 1),
                                      team = gsub("Women's","",team),
                                      .keep="unused")

```


```{r, include=FALSE,eval=T}

# Events for the games of the WC2023
WC2023_dataframe <- free_allevents(MatchesDF = Matches, Parallel = T)
WC2023_dataframe = allclean(WC2023_dataframe)

```


```{r, include=F, eval=F}

# Lire le fichier CSV dans R
#write.csv(WC2023_dataframe, "WC2023_dataframe.csv", row.names = FALSE)

WC2023_dataframe <- read.csv("WC2023_dataframe.csv")

```



```{r,, include=FALSE}
WC2023_dataframe <- WC2023_dataframe %>%
  mutate(team.name = str_remove(team.name, " Women's"))

WC2023_dataframe <- WC2023_dataframe %>%
  mutate(possession_team.name = str_remove(possession_team.name, " Women's"))
```


```{r, include=FALSE}
#For descriptive analysis and visualization we need to have some clearer way to colored datas, so we create groups depending on which stage of the competitions teams have lost
Semifinals <- AllMatches %>% filter(competition_stage.name=="Semi-finals")
pays_semifinals <- subset(Semifinals, select = c(home_team.home_team_name,away_team.away_team_name))
pays_semifinals <- melt(pays_semifinals) %>% distinct() %>% as.list() %>% unlist() %>% sort() #select all teams
pays_semifinals_list <- melt(pays_semifinals) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #select all teams
#view(pays_semifinals)

Quarterfinals <- AllMatches %>% filter(competition_stage.name=="Quarter-finals")
pays_quarterfinals <- subset(Quarterfinals, select = c(home_team.home_team_name,away_team.away_team_name))
pays_quarterfinals <- melt(pays_quarterfinals) %>% distinct() %>% as.list() %>% unlist() %>% sort() #select all teams
pays_quarterfinals <- setdiff(pays_quarterfinals, pays_semifinals)
pays_quarterfinals_list <- melt(pays_quarterfinals) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #select all teams
#view(pays_quarterfinals)

Round_of_16 <- AllMatches %>% filter(competition_stage.name=="Round of 16")
pays_r16 <- subset(Round_of_16, select = c(home_team.home_team_name,away_team.away_team_name))
pays_r16 <- melt(pays_r16) %>% distinct() %>% as.list() %>% unlist() %>% sort() #select all teams
pays_r16 <- setdiff(pays_r16, union(pays_quarterfinals, pays_semifinals))
pays_r16_list <- melt(pays_r16) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #select all teams
#view(pays_r16)

Group_Stage <- AllMatches %>% filter(competition_stage.name=="Group Stage")
pays_gs <- subset(Group_Stage, select = c(home_team.home_team_name,away_team.away_team_name))
pays_gs <- melt(pays_gs) %>% distinct() %>% as.list() %>% unlist() %>% sort() #select all teams
pays_r8 <- union(pays_quarterfinals, pays_semifinals)
pays_gs <- setdiff(pays_gs, union(pays_r16, pays_r8))
pays_gs_list <- melt(pays_gs) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #select all teams
#view(pays_gs)

# Créer un data frame pour l'échelle de couleur
color_scale_data_GS <- data.frame(team = pays_gs, group = "8-GS")
color_scale_data_R16 <- data.frame(team = pays_r16, group = "7-R16")
color_scale_data_Quarter <- data.frame(team = pays_quarterfinals, group = "6-Quarterfinals")
color_scale_data_Semi <- data.frame(team = pays_semifinals, group = "5-Semifinals")
color_scale_data_4 <- data.frame(team = "Australia", group = "4th")
color_scale_data_3 <- data.frame(team = "Sweden", group = "3rd")
color_scale_data_2 <- data.frame(team = "England", group = "2nd")
color_scale_data_1 <- data.frame(team = "Spain", group = "1st")
color_scale_data <- bind_rows(color_scale_data_GS, 
                              color_scale_data_R16, 
                              color_scale_data_Quarter, 
                              color_scale_data_Semi)
color_scale_2 <- bind_rows(color_scale_data_GS, 
                              color_scale_data_R16, 
                              color_scale_data_Quarter, 
                              color_scale_data_4, color_scale_data_3, color_scale_data_2, color_scale_data_1)

#view(color_scale_data)
```

# Descriptive data analysis 

The package StatsbombR provides the data from 71 national and international competitions, for a total of over 3000 matches. For the sake of this project, we narrow our scope down to the most recent competition available : The FIFA Women's World Cup 2023 and its 64 matches. We begin by interpreting the different variables of this large data set.\

The full data set contains 183 variables for analysis, with the majority having a significant proportion of missing values, as they were used to track very specific patterns of play. As an example, the parameter "goalkeeper.shot_saved_to_post" is attributed "True" only if the goalkeeper saved a shot from going inside the goal, by deflecting it onto a post.\


```{r, include=FALSE}
create_Pitch <- function(opposing=TRUE, grass_colour = "#F9F9F9", line_colour = "#8F8F8F", background_colour = "#F9F9F9", goal_colour = "#000000", goaltype = "line", middlethird = FALSE, BasicFeatures = FALSE, JdeP = FALSE, arcs = TRUE, padding = 5, Heatmap = FALSE, df_heatmap){

  library(ggplot2)
  ## set theme for blank pitch
  theme_blankPitch = function(size=12) {
    theme(
      #axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      #axis.ticks.y=element_text(size=size),
      #   axis.ticks=element_blank(),
      axis.ticks.length=unit(0, "lines"),
      #axis.ticks.margin=unit(0, "lines"),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      legend.background=element_rect(fill=background_colour, colour=NA),
      legend.key=element_rect(colour=background_colour,fill=background_colour),
      legend.key.size=unit(1.2, "lines"),
      legend.text=element_text(size=size),
      legend.title=element_text(size=size, face="bold",hjust=0),
      strip.background = element_rect(colour = background_colour, fill = background_colour, linewidth = .5),
      panel.background=element_rect(fill=background_colour,colour=background_colour),
      #       panel.border=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      panel.spacing=element_blank(),
      plot.background=element_blank(),
      plot.margin=unit(c(0, 0, 0, 0), "lines"),
      plot.title=element_text(size=size*1.2),
      strip.text.y=element_text(colour=background_colour,size=size,angle=270),
      strip.text.x=element_text(size=size*1))}

  ymin <- 0 # minimum width
  ymax <- 80 # maximum width
  if (opposing==TRUE){
  xmin <- 80 # minimum length
  }else{
    xmin <-0
    }
  xmax <- 120 # maximum length

  # Defining features along the length
  boxEdgeDef <- 18
  boxEdgeOff <- 102
  halfwayline <- 60
  sixYardDef <- 6
  sixYardOff <- 114
  penSpotDef <- 12
  penSpotOff <- 108

  # Defining features along the width
  boxEdgeLeft <- 18
  boxEdgeRight <- 62
  sixYardLeft <- 30
  sixYardRight <- 50
  goalPostLeft <- 36
  goalPostRight <- 44
  CentreSpot <- 40

  # other dimensions
  centreCirle_d <- 20

  ## define the circle function
  circleFun <- function(center = c(0,0),diameter = 1, npoints = 100){
    r = diameter / 2
    tt <- seq(0,2*pi,length.out = npoints)
    xx <- center[1] + r * cos(tt)
    yy <- center[2] + r * sin(tt)
    return(data.frame(x = xx, y = yy))
  }

  #### create center circle ####
  center_circle <- circleFun(c(halfwayline,CentreSpot),centreCirle_d,npoints = 100)

  ### FIRST STAGE
  ## initiate the plot, set some boundries to the plot
  if (Heatmap == TRUE){
    p <- ggplot(df_heatmap, aes(x = x, y = y, fill = xg)) + 
      xlim(c(xmin-padding,xmax+padding)) + ylim(c(ymin-padding,ymax+padding)) +   
      theme_blankPitch() + 
      geom_tile() + 
      scale_fill_viridis(option = "A")
  }else{
    p <- ggplot() + xlim(c(xmin-padding,xmax+padding)) + ylim(c(ymin-padding,ymax+padding)) + theme_blankPitch()
  }# add the theme

  ### ADD MIDDLE THIRD SHADING
  if(middlethird == TRUE){
  p <- p + geom_rect(aes(xmin=(xmax/3*1), xmax=(xmax/3*2), ymin=ymin, ymax=ymax), colour = NA, fill = "black", alpha = 0.10)
  }else{}

  if(BasicFeatures == TRUE){
  p <- p +
  # add the base rectangle of the pitch
  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill = NA, colour = line_colour) +
  # add the 18 yard box defensive
  geom_rect(aes(xmin=xmin, xmax=boxEdgeDef, ymin=boxEdgeLeft, ymax=boxEdgeRight), fill = grass_colour, colour = line_colour) +
  # add the 18 yard box offensive
  geom_rect(aes(xmin=boxEdgeOff, xmax=xmax, ymin=boxEdgeLeft, ymax=boxEdgeRight), fill = grass_colour, colour = line_colour) +
  # add halway line
  geom_segment(aes(x = halfwayline, y = ymin, xend = halfwayline, yend = ymax),colour = line_colour)
  arcs = FALSE
  }else{
        ## initiate the plot, set some boundries to the plot
  p <- p +
  # add the base rectangle of the pitch
  geom_rect(aes(xmin=xmin, xmax=xmax, ymin=ymin, ymax=ymax), fill = NA, colour = line_colour) +
  # add the 18 yard box defensive
  geom_rect(aes(xmin=xmin, xmax=boxEdgeDef, ymin=boxEdgeLeft, ymax=boxEdgeRight), fill = grass_colour, colour = line_colour) +
  # add the 18 yard box offensive
  geom_rect(aes(xmin=boxEdgeOff, xmax=xmax, ymin=boxEdgeLeft, ymax=boxEdgeRight), fill = grass_colour, colour = line_colour) +
  # add halway line
  geom_segment(aes(x = halfwayline, y = ymin, xend = halfwayline, yend = ymax),colour = line_colour) +
  # add the six yard box Defensive
  geom_rect(aes(xmin=xmin, xmax=sixYardDef, ymin=sixYardLeft, ymax=sixYardRight), fill = grass_colour, colour = line_colour)  +
  # add the six yard box offensive
  geom_rect(aes(xmin=sixYardOff, xmax=xmax, ymin=sixYardLeft, ymax=sixYardRight), fill = grass_colour, colour = line_colour) +
  # add centre circle
  #geom_path(data=center_circle, aes(x=x,y=y), colour = line_colour) +
    # add penalty spot left
  geom_point(aes(x = penSpotDef , y = CentreSpot), colour = line_colour, size = 0.75) +
  # add penalty spot right
  geom_point(aes(x = penSpotOff , y = CentreSpot), colour = line_colour, size = 0.75) +
  # add centre spot
  geom_point(aes(x = halfwayline , y = CentreSpot), colour = line_colour, size = 0.75) }
  
  #### add goals depending on type

  ## LINE TYPE
  if(goaltype == "line"){
  p <- p +
  # add the goal Defensive
  geom_segment(aes(x = xmin, y = goalPostLeft, xend = xmin, yend = goalPostRight),colour = goal_colour, size = 1) +
  # add the goal offensive
  geom_segment(aes(x = xmax, y = goalPostLeft, xend = xmax, yend = goalPostRight),colour = goal_colour, size = 1)

  }else{}

  ## Barca Numbers TYPE
  if(goaltype == "barcanumbers"){
  p <- p +
  # add the goal Defensive
  geom_segment(aes(x = xmin - 0.75, y = goalPostLeft, xend = xmin - 0.75, yend = goalPostRight),colour = line_colour, size = 0.75) +
  # add the goal offensive
  geom_segment(aes(x = xmax + 0.75, y = goalPostLeft, xend = xmax + 0.75, yend = goalPostRight),colour = line_colour, size = 0.75)

  }else{}

  ## BOX TYPE
  if(goaltype == "box"){
  p <- p +
  # add the goal Defensive
  geom_rect(aes(xmin = xmin - 2 , ymin = goalPostLeft, xmax = xmin, ymax = goalPostRight), fill = grass_colour, colour = line_colour) +
  # add the goal offensive
  geom_rect(aes(xmin = xmax, ymin = goalPostLeft, xmax = xmax + 2, ymax = goalPostRight), fill = grass_colour, colour = line_colour)
  }else{}


  ## add J de P
  if(JdeP == TRUE){
  p <- p +
  # vertical tram lines
  geom_segment(aes(x = boxEdgeDef, y = boxEdgeLeft, xend = boxEdgeOff, yend = boxEdgeLeft), colour = "#941C07", alpha = 0.3) +
  geom_segment(aes(x = boxEdgeDef, y = boxEdgeRight, xend = boxEdgeOff, yend = boxEdgeRight), colour = "#941C07", alpha = 0.3) +
  geom_segment(aes(x = boxEdgeDef, y = CentreSpot - 10, xend = boxEdgeOff, yend = CentreSpot - 10), colour = "#941C07", alpha = 0.3) +
  geom_segment(aes(x = boxEdgeDef, y = CentreSpot + 10, xend = boxEdgeOff, yend = CentreSpot + 10), colour = "#941C07", alpha = 0.3) +
  # horizontal tram lines
  geom_segment(aes(x = boxEdgeDef, y = ymin, xend = boxEdgeDef, yend = ymax), colour = "#941C07", alpha = 0.3) +
  geom_segment(aes(x = boxEdgeOff, y = ymin, xend = boxEdgeOff, yend = ymax), colour = "#941C07", alpha = 0.3) +
  geom_segment(aes(x = (xmax/3*1), y = boxEdgeRight, xend = (xmax/3*1), yend = ymax), colour = "#941C07", alpha = 0.3) +
  geom_segment(aes(x = (xmax/3*1), y = boxEdgeLeft, xend = (xmax/3*1), yend = ymin), colour = "#941C07", alpha = 0.3) +
  geom_segment(aes(x = (xmax/3*2), y = boxEdgeRight, xend = (xmax/3*2), yend = ymax), colour = "#941C07", alpha = 0.3) +
  geom_segment(aes(x = (xmax/3*2), y = boxEdgeLeft, xend = (xmax/3*2), yend = ymin), colour = "#941C07", alpha = 0.3)
  # add the 18 yard box defensive
  #geom_rect(aes(xmin=xmin, xmax=boxEdgeDef, ymin=boxEdgeLeft, ymax=boxEdgeRight), fill = NA, colour = line_colour) +
  # add the 18 yard box offensive
  #geom_rect(aes(xmin=boxEdgeOff, xmax=xmax, ymin=boxEdgeLeft, ymax=boxEdgeRight), fill = NA, colour = line_colour)
  }else{}

  ## add J de P
  if(arcs == TRUE){
  p <- p +
  # vertical tram lines
  annotate("path",
           x = 12 + 10 * cos(seq(-0.3*pi, 0.3*pi, length.out = 30)),
           y = 40 + 10 * sin(seq(-0.3*pi, 0.3*pi, length.out = 30)),
           col = line_colour) +
  annotate("path",
           x = (120-12) - 10 * cos(seq(-0.3*pi, 0.3*pi, length.out = 30)),
           y = 40 + 10 * sin(seq(-0.3*pi, 0.3*pi, length.out = 30)),
           col = line_colour)
  }else{}

  return(p)

}

# This function will be useful to create a pitch background for plotting.
```

## Analysis of successful shots according to country

First, we will look at how the data is stored in the data set. We will do this by plotting the shots of a specific match; we chose Spain-England, which was the final match of the competition.

```{r, echo = F, eval = T, warning=FALSE, fig1 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig1}Visualization of the shots of Spain-England on the pitch",fig.height=4.5}
Visualization_df <- WC2023_dataframe %>% subset(match_id == 3906390) %>% filter(type.name == "Shot")

create_Pitch() + 
  geom_point(data = Visualization_df, aes(x=location.x, y=location.y)) + 
  geom_segment(data = Visualization_df, aes(x = location.x, y = location.y, xend = shot.end_location.x, yend = shot.end_location.y, color = player.name), show.legend = F)

#view(Visualization_df)
```

We can see that regardless of which team scores, the locations of the players are always tracked the same way : for a given team, the home goal is located at x = 0, which corresponds to the left of this graph, and the opposing goal is located at x = 120. This will allow us to directly use the provided variables, without further formatting the data.

Then, we took a look at the number of goals and shots in all matches for each team.\
Figure \ref{fig:fig1} shows a visualization of these results.\

```{r, include=FALSE}
#Number of goals and shots in all games for each team
shots_goals = WC2023_dataframe %>%
group_by(team.name) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 
```

From the above figure, we can already see a big disparity in team success. Some teams, like Vietnam and Haiti, didn't even manage to score a goal over the course of the competition, while Sweden and France were more successful in this aspect. However, simply displaying how much shots and goals a team made provides an incomplete understanding of team effectiveness, and is generally a flawed metric for comparison, as teams that made it further into the competition naturally scored more goals, and had more shots. 

```{r fig1 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig1}Diagram of the number of goals and shots in all matches for each team",fig.height=4.5}
#Let's make a graph
ggplot(shots_goals, aes(y = reorder(as.factor(team.name), goals))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of goals and shots in all games for each team",
       y = "Team",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green"))
```

Therefore, we calculated the percentage of shots leading to a goal for every team, and compared them in Figure \ref{fig:fig2}.

From this graph, we are able to identify that Sweden was the most efficient team in scoring by over 2.5%. This is of course very biased, as Sweden finished third in the tournament, only losing one match throughout the entirety of the competition. Keeping that in mind, it is surprising that Spain scores this low on the graph, considering they won the World Cup. It could be that Spain, despite their evident success, was not a very efficient team, or that they had a different playing style than other teams.

```{r fig2 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig2}Diagram of the percentage of shots leading to a goal",fig.height=4,fig.align='center'}

ggplot(shots_goals, aes(y = reorder(as.factor(team.name), goals/shots*100))) +
  geom_bar(aes(x = goals/shots*100), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Percentage of shots leading to a goal",
       y = "Team",
       x = "Percentage")
```

\vspace{3pt}

Next, we wanted to realize the above analyses for singular matches. We chose the four matches played by team France, and created the same graphs for each of their matches.
Figure \ref{fig:fig3} shows the results.\

As we can see, there were a huge number of goals in two games : Panama France and Australia France. The first match ended on a exceptional 3-6 score for France, but the former ended on a nil-nil. However, teams went to penalties, and they scored a total of 13, making this observation flawed.

```{r,include=F}
shots_goals_all_matches = WC2023_dataframe %>%
  group_by(match_id) %>%
  reframe(team_name = unique(team.name),
            shots = sum(type.name=="Shot", na.rm = TRUE),
            goals = sum(shot.outcome.name=="Goal", na.rm = TRUE))

#Number of goals and shots in each match for both teams
#shots_goals_all_matches = WC2023_dataframe %>%
#group_by(match_id) %>% 
#summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
#goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 

```

```{r,include=F}
#We display ids of all France matches
Id_France = WC2023_dataframe%>% filter(team.name=="France")
match_ids_france=unique(Id_France$match_id)
print(match_ids_france)
```


```{r,include=F}
#For each French match, the total number of shots and goals for the 2 teams (France+adversary) is displayed.

france_goals <- shots_goals_all_matches %>%
  filter(match_id %in% match_ids_france) %>%
  subset(team_name!="France")

print(france_goals)
```


```{r,include=F}
ggplot(france_goals, aes(y = team_name)) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of shots and goals for each French match",
       y = "France's opposing team",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green"))
  
```


```{r,include=F}
ggplot(france_goals, aes(y = reorder(as.factor(match_id), goals/shots*100))) +
  geom_bar(aes(x = goals/shots*100), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Percentage of shots on goal in France matches",
       y = "Match",
       x = "Percentage")

#Il faut ordonner par date pour avoir l'évolution de la stratégie au cours de la compétition


```

```{r,include=F}

#We want to know which goals come from France
shots_goals_France_all_matches = WC2023_dataframe %>%
filter(team.name=="France") %>%
group_by(match_id) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 

print (shots_goals_France_all_matches)

```



```{r fig3 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig3}Diagram of the number of shots and goals for each French and percentage",fig.height=2,fig.align='center'}

g1=ggplot(shots_goals_France_all_matches, aes(y = as.factor(match_id))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Numbers for France's matches",
       y = "Match_id",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green")) 


g2=ggplot(shots_goals_France_all_matches, aes(y = as.factor(match_id))) +
  geom_bar(aes(x = goals/shots*100), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Percentage in France's matches",
       y = "Match",
       x = "Percentage") 

#Order from bottom to top: 1st match to last
grid.arrange(g1,g2,ncol=2)
```

## Analysis of successful shots according to different variables  

We first look at the different types of shots in Figure \ref{fig:fig3}. Four different types of shots were differentiated in the data set : "Open Play", "Penalty", "Free Kick" as well as "Corner." A shot was deemed as being "Open Play" if it was taken during regular actions of the game. The "Corner" label only applies to a single shot made by Ireland's Katie McCabe, and went in. This is something to keep in mind, as it is sure to skew our future models. Other labels are self-explanatory.\

```{r,include=F}
#Number of goals and shots in all matches by type of shot
shots_goals_formation = WC2023_dataframe %>%
group_by(shot.type.name) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 

shots_goals_formation=shots_goals_formation[-c(5),]
```


```{r fig4 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig4}Diagram of the number of shots and goals for each type of shot",fig.height=2,fig.align='center'}
ggplot(shots_goals_formation, aes(y = reorder(as.factor(shot.type.name), goals))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of shots and goals for each type of shot",
       y = "type",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green"))
```


Next, we look at the different techniques used by players : how much are being kept a track in the data set, how much were each of them used, and which one produced the most goals.\

Figure \ref{fig:fig4} shows that data set contains seven types of shots, although only three are consistently being used, that is : "Normal", "Half Volley" and "Volley". Naturally, the "Normal" shot was the most popular, and hence yielded the mos t goals. A "Lob" was very infrequently done, but could prove effective in the right situation : it seems a good portion of these shots went in.

```{r,include=F}
#Number of shots and goals for each technique of shot
shots_goals_technique = WC2023_dataframe %>%
group_by(shot.technique.name) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 

#On enlève la ligne NA

shots_goals_technique=shots_goals_technique[-c(8),]
```


```{r fig5 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig5}Diagram of the number of shots and goals for each technique of shot",fig.height=2,fig.pos="h",fig.align='center'}

ggplot(shots_goals_technique, aes(y = reorder(as.factor(shot.technique.name), goals))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of shots and goals for each technique of shot",
       y = "Technique of shot",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green"))
```

Similarly, we visualize in Figure \ref{fig:fig5} the different body parts used in shooting.

Unsurprisingly, the right foot was most commonly used, and it seems every body part was equally as effective in scoring, apart from the "Other" body zone.

```{r,include=F}
#Number of goals and shots over all matches according to body zone used
shots_goals_part = WC2023_dataframe %>%
group_by(shot.body_part.name) %>% 
summarise(shots = sum(type.name=="Shot", na.rm = TRUE),
goals = sum(shot.outcome.name=="Goal", na.rm = TRUE)) 
shots_goals_part=shots_goals_part[-c(5),]

```

```{r fig6 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig6}Diagram of the number of goals and shots according to body zone used",fig.height=2,fig.pos="h"}

ggplot(shots_goals_part, aes(y = reorder(as.factor(shot.body_part.name), goals))) +
  geom_bar(aes(x = shots, fill = "shots"), stat = "identity", position = "dodge", width = 0.8) +
  geom_bar(aes(x = goals, fill = "goals"), stat = "identity", position = "dodge", width = 0.8) +
  labs(title = "Number of goals and shots according to body zone used",
       y = "Body zone",
       x = "Number") +
  scale_fill_manual(values = c("shots" = "brown", "goals" = "green")) 
```


## Analysis of teams and players scoring ratio

In this part we want to describe the number of goals scored by players and teams by using the goal ratio, so the number of goals scored on the number of totals shots during the events.\

### Players scoring ratio

Firstly, we looked at players, this may lead to find outliers so players which have performed a lot of shots but only few goals, and the opposite players which attempted only few shots but scored on a lot of them. This is a very basic way to start taking an interest in performance of players during this Word Cup.

Figure \ref{fig:fig7} shows a visualization of the goal ratio on the total shots performed by players which have performed at least 1 shots during the World Cup.\


```{r,include=F}
df_shots_total <- aggregate(x = Shots$outcome, by = list(Shots$player), FUN = length)
df_shots_succeed <- aggregate(x = Shots$outcome, by = list(Shots$player), FUN = sum)

df <- merge(df_shots_succeed, df_shots_total, by=c("Group.1")) %>% mutate(player = Group.1,
                                                                               completed_shots = x.x,
                                                                               total_shots = x.y,
                                                                               .keep="unused")

df1 <- subset(Shots, selec = c(player, team))
#view(df1)

df <- merge(df1, df, by=c("player"))
#view(df)
df <- df %>% distinct()

df <- df %>% mutate(shots_completion_ratio = completed_shots/total_shots,
                    total_shots = total_shots,
                    total_goals = completed_shots,
                    .keep = "unused")

print(length(df$player))
#view(df)

df <- df[df$total_shots>=1,] # Removing players that attempted less than 2 shots

print(length(df$player)) # 398 players attempted at least 1 shots in the WC

#We now have a df containing every player that shot at least once in the World cup, associated with how many shots they attempted, and their ratio of successful shots

#pays = c("France", "England", "Spain", "Switzerland") #Select teams we wish to observe
pays <- melt(df$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #Select all teams

df_player <- df[df$team %in% pays,]

df$groupv2 <- ifelse(df$team %in% pays_gs, "8-GS",
                              ifelse(df$team %in% pays_r16, "7-R16",
                                     ifelse(df$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(df$team == "Australia", "4th",
                                                   ifelse(df$team == "Sweden", "3rd",
                                                          ifelse(df$team == "England", "2nd",
                                                                 ifelse(df$team=="Spain","1st","NA")))))))

mean_shots_completion_ratio = mean(df$shots_completion_ratio)
mean_shots_nb = mean(df$total_shots)
mean_goals_nb = mean(df$total_goals)


df[nrow(df) + 1,] <- list("Mean Player", "Mean Player", mean_shots_nb, mean_shots_completion_ratio, mean_goals_nb, "Mean Player")
#view(df)
```


```{r fig7 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig7}Goal ratio on the number of totals shots for players",fig.height=4.5,fig.align='center'}
p <- ggplot(df, aes(total_shots, shots_completion_ratio, label = paste("player =", player, ", team =", team ), color = groupv2)) + geom_point() +  scale_color_viridis(discrete=TRUE) + geom_point(data = df[df$player == "Mean Player", ], aes(total_shots, shots_completion_ratio), shape = 4, color = "red4", size = 2) + labs(x = "Total Number of Shots", y = "True Goal's Ratio", fill = "World Cup's Stages", color = "World Cup's Stages")

# Display “Mean Player” with a red cross
#p <- p + geom_point(data = df[df$player == "Mean Player", ], aes(total_shots, shots_completion_ratio), shape = 4, color = "red4", size = 2)

ggplotly(p)
p
```

Some information on this graph: the color of the dots represents, as indicated in the legend, at what stage of the competition the player's team finished, so the more the dots tend towards purple, the more matches the players played and therefore went further in the competition.
The graph shows a clear trend: the more players have shot, the lower their ratio. Even so, the players on the best teams, those who finished in the best positions, often have a higher ratio than those on teams who lost earlier in the tournament.
We also notice that the average (the red cross) is quite low compared to the points we see on the graph, intuitively we'd probably have placed it around 10 shots and 0.25 goal ratio. But then we realized that many points, especially those at the bottom of the graph with 0 goals scored, are superimposed, so there are many players who shot without scoring, which is fairly consistent with the match statistics of around ten shots per team for an average of between 1 and 4 goals per match.
Now we miss an important information on this graphics. Indeed we can see that the most players attempted shots the less their ratio of successful shots is high, but we don't take into account the difficulty of the shots perfomed by the players. Indeed a penalty is intuitively easier to score than an open shot far for the goals and under the pressure of defenders. It's something we already saw in the Figure \ref{fig:fig4}, that goals on penalty occurred more often even though that there are less frequent in games.

So, we ask ourselves base on which criteria we could try to implement the notion of shot's difficulty, and we find that the notion of Expected Goals is what we were looking for and that the Statsbomb dataset provide it with every shots attempt.

The "Expected Goal", often named "xG" is the probability to score given a lot of datas on the shot, for example it can be the shhoter and the goal position, the fact that the striker is under pressure or not, with which foot he is attempting this shot and various differents variables. This probability is computed by the xG model of Statsbomb which is probably created on a large amount of data that they were able to collect.

So this notion of xG is perfect to implement the notion of shots difficulty. We can compute the mean of the xG of every shots that one player attempted to see if the player had easy or hard shots to perform.

This his what the Figure \ref{fig:fig8} is showing true goal ratio on the Statsbomb xG ratio for every player which attempted at least 1 shots during the World Cup.\

```{r,include=F}
Shots_exp <- subset(Shots_All, select = c(shot.statsbomb_xg,team, player.name))

Shots_exp <- Shots_exp  %>%  mutate(player = player.name,
                                      xg = shot.statsbomb_xg,
                                      team = gsub("Women's","",team),
                                      .keep="unused")

#view(Shots_exp)

df_shots_exp_succeed <- aggregate(x = Shots_exp$xg, by = list(Shots_exp$player), FUN = sum)
#df_shots_exp_succeed <- df_shots_exp_succeed %>% mutate(player=Group.1, 
#                               df_shots_exp_ratio=df_shots_exp_succeed/df_shots_exp_total)

#view(df_shots_exp_succeed)
colnames(df_shots_exp_succeed) <- c("player", "xG")
#view(df)
#view(df_shots_exp_succeed)
df2 <- merge(df, df_shots_exp_succeed, by=c("player"))
df2 <- df2 %>% mutate(xG_ratio = df_shots_exp_succeed$xG/total_shots)
#,                  .keep = "unused")

#view(df2)

print(length(df2$player))
#view(df2)

df2 <- df2[df2$total_shots>=1,] # Removing players that attempted less than 1 shots

print(length(df2$player)) # 398 players attempted at least 1 shots in the WC

#We now have a df containing every player that shots at least once in the World cup, associated with their xg

```

```{r,include=F}
pays = c("France", "England", "Spain", "Switzerland") #Select teams we wish to observe
pays <- melt(df$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #Select all teams

df2_player <- df2[df2$team %in% pays,]



df2_player$groupv2 <- ifelse(df2_player$team %in% pays_gs, "8-GS",
                              ifelse(df2_player$team %in% pays_r16, "7-R16",
                                     ifelse(df2_player$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(df2_player$team == "Australia", "4th",
                                                   ifelse(df2_player$team == "Sweden", "3rd",
                                                          ifelse(df2_player$team == "England", "2nd",
                                                                 ifelse(df2_player$team=="Spain","1st","NA")))))))


mean_shots_completion_ratio = mean_goals_nb/mean_shots_nb
mean_shots_nb = mean(df2_player$total_shots)
mean_goals_nb = mean(df2_player$total_goals)
mean_shots_xG =  mean(df2_player$xG)
mean_shots_xG_ratio = mean_shots_xG/mean_shots_nb

df2_player[nrow(df2_player) + 1,] <- list("Mean Player", "Mean Player", mean_shots_nb, mean_shots_completion_ratio, mean_goals_nb, "Mean Player", mean_shots_xG, mean_shots_xG_ratio)



#view(df2_player)
```


```{r fig8 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig8}Goal ratio on the Expected Goal ratio (Statsbomb model) for players",fig.height=4,fig.align='center'}
p2 <- ggplot(df2_player, aes(xG_ratio, shots_completion_ratio, size = total_shots, label = player, colour=groupv2)) + 
    geom_point(shape=21) + geom_abline(intercept = 0, slope = 1, linetype = "dashed") + scale_color_viridis(discrete=TRUE) + geom_point(data = df2_player[df2_player$team == "Mean Player", ], aes(xG_ratio, shots_completion_ratio, size = total_shots), shape = 4, color = "red4", stroke = 1) + labs(x = "Expected Goal's ratio", y = "True Goal's Ratio", fill = "World Cup's Stages", color = "World Cup's Stages")

# Display “Mean Player” with a red cross

#p2 <- p2 + geom_point(data = df2_player[df2_player$team == "Mean Player", ], aes(xG_ratio, shots_completion_ratio, size = total_shots), shape = 4, color = "red4", stroke = 1)

ggplotly(p2)
p2
```

This graph compares the number of goals predicted by Statsbomb with the number of goals actually scored. It shows whether players scored more or fewer goals than predicted by the Statsbomb model. We find the same biases as in the graph above: players who scored on their only shot are obviously better performers, but this is no guarantee of performance, as we can't assess the notion of regularity. The same applies to all players represented by the smaller circles.
As none of the axes in this graph display the notion of number of shots taken, we are obliged to add this notion by modifying the size of the circle representing the player. This reinforces the caution of interpretation in relation to the number of matches played and how often they had shot during the game.

Now if we interpret this graph, we can see from the colors that the majority of players who have reached the quarter-finals or beyond are still around the dotted line that symbolizes player performance. This line, with the equation y=x, symbolizes the expected performance of the players. Thus, we can deduce that players from teams that made it to the quarter-finals and beyond underperformed less often and even over-performed less, since they took more shots and so their data is less biased by a successful difficult shot.\


### Teams scoring ratio

Secondly, we decided to look at teams, we are running the same analysis as for the players but this could lead to other analysis. Indeed it's reducing the number of outliers, because we are conduction an analysis on all team players that have shot at least once during the tournament. Indeed if a midfielder player has shot only once and scored and a striker has shot 10 times for only 2 goals this bias of the midfielder has less impact on her team. So, with this team analysis we could say overall which team underperform inter of goals, and try to find if it's correlated to the team who lost earliest during the competition.

Figure \ref{fig:fig9} shows a visualization of the goal ratio on the total shots performed by teams which have performed at least 1 shots during the World Cup.\

```{r,include=F}
df_shots_total <- aggregate(x = Shots$outcome, by = list(Shots$player), FUN = length)
df_shots_succeed <- aggregate(x = Shots$outcome, by = list(Shots$player), FUN = sum)

df <- merge(df_shots_succeed, df_shots_total, by=c("Group.1")) %>% mutate(player = Group.1,
                                                                               completed_shots = x.x,
                                                                               total_shots = x.y,
                                                                               .keep="unused")

df1 <- subset(Shots, selec = c(player, team))
#view(df1)

df <- merge(df1, df, by=c("player"))
#view(df)
df <- df %>% distinct()

df <- df %>% mutate(shots_completion_ratio = completed_shots/total_shots,
                    total_shots = total_shots,
                    total_goals = completed_shots,
                    .keep = "unused")

print(length(df$player))
#view(df)

df <- df[df$total_shots>=1,] # Removing players that attempted less than 1 shots

print(length(df$player)) # 398 players attempted at least 1 shots in the WC

#We now have a df containing every player that shots at least once in the World cup, associated with how many shots they attempted, and their ratio of successful shots

#pays = c("France", "England", "Spain", "Switzerland") #Countries we wish to observe
#pays = c("Switzerland")
pays <- melt(df$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #select all teams

#view(pays)

df_pays <- df[df$team %in% pays,]

for (i in 0:length(pays))
  df_pays[nrow(df_pays) + 1,] <- list(paste("",pays[i]), paste("", pays[i]), 
                                      sum(df_pays$total_shots[df_pays$team == pays[i]]), 
                                      0,
                                      sum(df_pays$total_goals[df_pays$team == pays[i]]))

df_pays <- subset(df_pays, select = -c(player))

n <- length(pays)

#view(df_pays)

dftail <- df_pays[(nrow(df_pays)-(n - 1)):nrow(df_pays),] %>% mutate(team = gsub(" ","",team),
                                                                     shots_completion_ratio = total_goals/total_shots,
                                                                     total_shots = total_shots,
                                                                     total_goals = total_goals,
                                                                     .keep = "unused")

#pays_gs <- pays_gs %>% mutate(team=gsub(" ","",x))

dftail$groupv2 <- ifelse(dftail$team %in% pays_gs, "8-GS",
                              ifelse(dftail$team %in% pays_r16, "7-R16",
                                     ifelse(dftail$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(dftail$team == "Australia", "4th",
                                                   ifelse(dftail$team == "Sweden", "3rd",
                                                          ifelse(dftail$team == "England", "2nd",
                                                                 ifelse(dftail$team=="Spain","1st","NA")))))))

mean_shots_nb_pays = mean(dftail$total_shots)
mean_goals_nb_pays = mean(dftail$total_goals)
mean_shots_completion_ratio_pays = mean_goals_nb_pays/mean_shots_nb_pays



dftail[nrow(dftail) + 1,] <- list("Mean Country", mean_shots_nb_pays, mean_shots_completion_ratio_pays, mean_goals_nb_pays, "Mean Country")

#view(dftail)
```


```{r fig9 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig9}Goal ratio on the number of total shots for teams",fig.width=8, fig.asp=0.8,fig.align='center'}
p1 <- ggplot(dftail, aes(total_shots, shots_completion_ratio, label=team, color = groupv2)) + geom_text(aes(fontface=2)) + scale_color_viridis(discrete=TRUE) + geom_point(data = dftail[dftail$team == "Mean Country", ], aes(total_shots, shots_completion_ratio), shape = 4, color = "red4", size = 4) + labs(x = "Total Number of Shots", y = "True Goal's Ratio", fill = "World Cup's Stages", color = "World Cup's Stages")

# Display “Mean Country” with a red cross
#p1 <- p1 + geom_point(data = dftail[dftail$team == "Mean Country", ], aes(total_shots, shots_completion_ratio), shape = 4, color = "red4", size = 4)

ggplotly(p1)
p1
```
In this graph, countries are represented by dots with their name, and the color always indicates the stage of the competition they have reached. There's a clear separation between the bottom left of the image, representing countries with few shots and few goals, with teams that didn't make it beyond the quarter-finals, and the top right with the 8 best teams, excluding Colombia.

We can also see that Spain, winner of the tournament, shot much more than its rivals but scored less, so this could reflect a strong domination during the matches, but less good finishing. It would be interesting to see if the difficulty of the shots could explain the lower accuracy, or if there's another interpretation.

Other outliers were Sweden, who finished 3rd and had the best ratio of successful shots. And Vietnam and Haiti, who didn't score a single goal during the competition.



As for the players, we want to implement the Expected Goal data, so we can implement and analyze the difficulty of shots performed by teams.

This his what the Figure \ref{fig:fig10} is showing true goal ratio on the Statsbomb xG ratio for every teams during the World Cup.\
```{r,include=F}
Shots_exp <- subset(Shots_All, select = c(shot.statsbomb_xg,team, player.name))

Shots_exp <- Shots_exp  %>%  mutate(player = player.name,
                                      xg = shot.statsbomb_xg,
                                      team = gsub("Women's","",team),
                                      .keep="unused")

#view(Shots_exp)

df_shots_exp_succeed <- aggregate(x = Shots_exp$xg, by = list(Shots_exp$player), FUN = sum)
#df_shots_exp_succeed <- df_shots_exp_succeed %>% mutate(player=Group.1, 
#                               df_shots_exp_ratio=df_shots_exp_succeed/df_shots_exp_total)

#view(df_shots_exp_succeed)
colnames(df_shots_exp_succeed) <- c("player", "xG")
#view(df)
#view(df_shots_exp_succeed)
df2 <- merge(df, df_shots_exp_succeed, by=c("player"))
df2 <- df2 %>% mutate(xG_ratio = df_shots_exp_succeed$xG/total_shots)
#,                  .keep = "unused")

#view(df2)

print(length(df2$player))
#view(df2)

df2 <- df2[df2$total_shots>=1,] # Removing players that attempted less than 1 shots

print(length(df2$player)) # 398 players attempted at least 1 shots in the WC

#We now have a df containing every player that shots at least once in the World cup, associated with their xg

#pays2 = c("France", "England", "Spain", "Switzerland") #Countries we wish to observe
pays2 <- melt(df2$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #select all teams

#view(pays2)

df2_pays <- df2[df2$team %in% pays2,]
#view(df2_pays)

for (i in 0:length(pays2))
  df2_pays[nrow(df2_pays) + 1, ] <- list(paste("",pays2[i]), paste("", pays2[i]),
                                        sum(df2_pays$total_shots[df2_pays$team == pays2[i]]),
                                        sum(df2_pays$shots_completion_ratio[df2_pays$team == pays2[i]]),
                                        sum(df2_pays$total_goals[df2_pays$team == pays2[i]]),
                                        sum(df2_pays$xG[df2_pays$team == pays2[i]]),
                                        0)
  
df2_pays <- subset(df2_pays, select = -c(player))

#view(df2_pays)
n <- length(pays2)

df2tail <- df2_pays[(nrow(df2_pays)-(n - 1)):nrow(df2_pays),]  %>% mutate(team = gsub(" ","",team),
                                                                      shots_completion_ratio = total_goals/total_shots,
                                                                     total_shots = total_shots,
                                                                     total_goals = total_goals,
                                                                     xG = xG,
                                                                     xG_ratio = xG/total_shots,
                                                                     .keep = "unused")

#view(df2tail)

df2tail$groupv2 <- ifelse(df2tail$team %in% pays_gs, "8-GS",
                              ifelse(df2tail$team %in% pays_r16, "7-R16",
                                     ifelse(df2tail$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(df2tail$team == "Australia", "4th",
                                                   ifelse(df2tail$team == "Sweden", "3rd",
                                                          ifelse(df2tail$team == "England", "2nd",
                                                                 ifelse(df2tail$team=="Spain","1st","NA")))))))

mean_shots_nb_pays = mean(df2tail$total_shots)
mean_goals_nb_pays = mean(df2tail$total_goals)
mean_shots_completion_ratio_pays = mean_goals_nb_pays/mean_shots_nb_pays
mean_shots_xG_pays = mean(df2tail$xG)
mean_shots_xG_ratio_pays = mean_shots_xG_pays/mean_shots_nb_pays

df2tail[nrow(df2tail) + 1,] <- list("Mean Country", mean_shots_nb_pays, mean_shots_completion_ratio_pays, mean_goals_nb_pays, mean_shots_xG_pays, mean_shots_xG_ratio_pays, "Mean Country")


#view(df2tail)
```


```{r fig10 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig10}Goal ratio on the Expected Goal ratio (Statsbomb model) for teams",fig.width=8, fig.asp=0.8,fig.align='center'}
p3 <- ggplot(df2tail, aes(xG_ratio, shots_completion_ratio, size = total_shots, label=team, color = groupv2)) + geom_text(aes(fontface=3)) + geom_abline(intercept = 0, slope = 1, linetype = "dashed") + scale_color_viridis(discrete=TRUE) + geom_point(data = df2tail[df2tail$team == "Mean Country", ], aes(xG_ratio, shots_completion_ratio, size = total_shots), shape = 4, color = "red4", stroke = 1) + labs(x = "Expected Goal's Ratio", y = "True Goal's Ratio", fill = "World Cup's Stages", color = "World Cup's Stages")

# Display “Mean Country” with a red cross
#p3 <- p3 + geom_point(data = df2tail[df2tail$team == "Mean Country", ], aes(xG_ratio, shots_completion_ratio, size = total_shots), shape = 4, color = "red4", stroke = 1)

ggplotly(p3)
p3
```

So, on this graph, we can complete our interpretations of the previous graph. Sweden, on the right-hand side of the graph, had the easiest shots on average according to Statsbomb's model, and succeeded as expected, or even a little better, as it lies slightly above the dotted line that marks the performance.

As for Spain, we can see that they had the hardest shots of all the teams that made it through to the quarter-finals, with the exception of Colombia, which explains why they had a greater number of shots with a fairly low ratio compared to their main rivals.

However, it is noticeable that the best teams have higher xGs, which means that their shots are more likely to succeed on average. This could mean that their forwards are putting themselves in better shooting situations, either because they are better or because the other teams' defenders are weaker. Thus, it could mean that the defense of the best teams is very strong compared to that of the teams eliminated earlier in the tournament, so these weaker teams had difficulty scoring.
Now that we have compared the true goal ratio to the xG Statsbomb ratio, we were wondering how exactly is one xG model working. So, we decided to create our own xG model.\

\clearpage

# Models analysis

We wanted to create our own xG model. To do that we developed different models, finding the most relevant variables to predict goals.\

```{r,include=F}
#We create a df with just the shots

shots = WC2023_dataframe %>%
 filter(type.name=="Shot")
#view(shots)
```

We run a logistic regression model: we want the output to be 0 or 1 depending on whether the shot turns into a goal.\

## First model : body part, technique, type of shot

The first model keeps the variables studied previously : body part, technique, type of shot. \

```{r,include=F}
df_model_1 <- dplyr::select(shots,shot.outcome.name,shot.body_part.name,shot.technique.name,shot.type.name)
df_model_1$shot.outcome.name <- ifelse(df_model_1$shot.outcome.name == "Goal", 1, 0)
```

```{r, include = F}
#Testing regression without interaction

glm.model_1<-glm(shot.outcome.name~ . ,data=df_model_1,family=binomial(link="logit"))
summary(glm.model_1)
```

```{r,include=F}
#We see a skewed shot.type.namePenalty because the corner modality is taken as reference and, given that there were no shots or goals during the competition, the model considers 100% success for the corner modality.


#We change the reference modality from “corner” to “penalty”.
df_model_1_modif=df_model_1
df_model_1_modif$shot.type.name <- relevel(as_factor(df_model_1_modif$shot.type.name), ref = "Penalty")

glm.model_1_modif <- glm(shot.outcome.name ~ . , data = df_model_1_modif, family = binomial(link = "logit"))
summary(glm.model_1_modif)
```

```{r,include=F}
#We can see that Open Play and Free Kick are less likely to result in a goal than a penalty kick. The corner remains distorted
```

```{r,include=F}
#R2 calculation:
R2_model1=1-(glm.model_1$deviance/glm.model_1$null.deviance)
print(R2_model1)
```
$R^2$ for the model without interaction is : \( `r R2_model1` \) \
```{r,include=F}
#Testing regression with interaction
glm.model_1_interac<-glm(shot.outcome.name~ .^2 ,data=df_model_1,family=binomial(link="logit"))
summary(glm.model_1_interac)
#A lot of NA because some interactions are impossible (e.g. volleyball restart on a penalty kick).
```


```{r,include=F}
#R2 calculation:
R2_model1_interac=1-(glm.model_1_interac$deviance/glm.model_1_interac$null.deviance)
print(R2_model1_interac)
```

$R^2$ for the model with interaction is : \( `r R2_model1_interac` \) .\

## Our target model : the expected goal variable

We now create a model composed of a single variable: the expected goal given in StatsBomb.\


```{r,include=F}
df_model_2 <- dplyr::select(shots,shot.statsbomb_xg,shot.outcome.name)
df_model_2$shot.outcome.name <- ifelse(df_model_2$shot.outcome.name == "Goal", 1, 0)

glm.model_2<-glm(shot.outcome.name~ . ,data=df_model_2,family=binomial(link="logit"))
summary(glm.model_2)
```

```{r,include=F}
R2_model2=1-(glm.model_2$deviance/glm.model_2$null.deviance)
print(R2_model2)
```

Our goal in creating the different models in this section is to find the most accurate model possible, which can have an $R^2$ close to this model (with only the expected goal as a variable), i.e. an $R^2$ close to : \( `r R2_model2` \).\

## Model 3 : Adding location.x and location.y

Any player on the field is assimilated as a moving point on a rectangle of size 80x120m. Its horizontal movement - that is, going from one goal post to another - is tracked by the variable location.x, while the vertical movement is associated to location.y. We will now add location.x and location.y to our previously adjusted model. \

```{r,include=F}
df_model_3 <- dplyr::select(shots,shot.outcome.name,shot.body_part.name,shot.technique.name,shot.type.name,location.x,location.y)

df_model_3$shot.outcome.name <- ifelse(df_model_3$shot.outcome.name == "Goal", 1, 0)
```

```{r,include=F}
#Model without interaction
glm.model_3<-glm(shot.outcome.name ~ . ,data=df_model_3,family=binomial(link="logit"))
summary(glm.model_3)
```


```{r,include=F}
R2_model_3=1-(glm.model_3$deviance/glm.model_3$null.deviance)
print(R2_model_3)
```
We test a regression without interaction, and obtain an $R^2$ of : \( `r R2_model_3` \) .\


```{r,include=F}
#Model with interaction

glm.model_3_interac<-glm(shot.outcome.name~ .^2 ,data=df_model_3,family=binomial(link="logit"))
summary(glm.model_3_interac)
```

```{r,include=F}
R2_model_3_interac=1-(glm.model_3_interac$deviance/glm.model_3_interac$null.deviance)
print(R2_model_3_interac)
```

With interactions, we get an $R^2$ of : \( `r R2_model_3_interac` \).\

In this model, we targeted the main variables to obtain a good model and an $R^2$ as close to 1 as possible.

### Significance of variables ?


We run several tests to see which variables are significant in the model. \

```{r,echo=F}
model_simp<-glm(shot.outcome.name~ (location.x +location.y + shot.body_part.name+shot.type.name)^2 ,data=df_model_3,family=binomial(link="logit"))
anova(model_simp,glm.model_3_interac,test="Chisq")
```

We see that we can remove the technique because $p-value$>0.05 so we can accept the sub-model with a 95% level.\

```{r,include=F}
summary(model_simp)
```


```{r,include=F}
R2_model_simp=1-(model_simp$deviance/model_simp$null.deviance)
print(R2_model_simp)
```

For this sub-model without the technique variable we obtain an $R^2$ of : \( `r R2_model_simp` \)

The $R^2$ is no greater than for model 3 with interactions: this is normal because the $R^2$ favors models with many variables.\
We should look at other variables such as AIC score, which is minimal for model 3 without interactions. \

```{r,include=F}
#We'll compare the fitted values (probability of being a goal), with the expected goals.
glm.model_3$fitted.values
#we have numbers between 0 and 1 so this is good.

```

### Comparison of norms

We now want to compare model 3 with and without interaction : the closer the 2-norm is to 0, the better the model. 


```{r,include=F}

#Display norm L2 for the model_3 without interaction

norm_L2_mod3 <- norm(glm.model_3$fitted.values - shots$shot.statsbomb_xg, type = "2")
print(norm_L2_mod3)

#Display norm L2 for the model_3 with interactions 

norm_L2_mod3withinteractions <- norm(glm.model_3_interac$fitted.values - shots$shot.statsbomb_xg, type = "2")
print(norm_L2_mod3withinteractions)


```

Norm L2 for the model_3 without interaction is equal to : \( `r norm_L2_mod3` \). \

The value for the model_3 with interactions is : \( `r norm_L2_mod3withinteractions` \).\

We find the same results as with the AIC criterion. This is consistent with the fact that $R^2$ favors models with many variables, so it's better to evaluate with AIC. We can conclude that the model 3 without interaction is best.

We do the same to compare model 1 with and without interaction.

```{r,include=F}

#Display norm L2 for the model_1 without interaction
norm_L2_mod1 <- norm(glm.model_1$fitted.values - shots$shot.statsbomb_xg, type = "2")
print(norm_L2_mod1)

#Display norm L2 for the model_1 with interactions
norm_L2_mod1withinteractions <- norm(glm.model_1_interac$fitted.values - shots$shot.statsbomb_xg, type = "2")
print(norm_L2_mod1withinteractions)

```

The L2 norms are respectively :  \( `r norm_L2_mod1` \) and \( `r norm_L2_mod1withinteractions` \). \

Both models are less accurate than the 3rd one.


## Model 4 : adding the under_pressure variable

We now create a new model like the model_3, but adding a variable : under_pressure.\

### Test for significance of single variables

First, we test the significance of this new variable.

```{r,echo=F}
df_model_4 <- dplyr::select(shots,shot.outcome.name,shot.body_part.name,shot.technique.name,shot.type.name,location.x,location.y,under_pressure)
df_model_4$shot.outcome.name <- ifelse(df_model_4$shot.outcome.name == "Goal", 1, 0)

#Replace NA with “false” in the under_pressure column
df_model_4$under_pressure <- ifelse(is.na(df_model_4$under_pressure), FALSE, df_model_4$under_pressure)

#Testing the model with only the under_pressure variable
glm.model_4 <- glm(shot.outcome.name ~ under_pressure, data = df_model_4, family = binomial(link = "logit"))
summary(glm.model_4)
```


We see that \( p_{\text{value}} < 0.05 \), so we reject $H_0$ : playing under pressure is significant. \

Estimated coefficients are negative, so playing under pressure reduces the probability of scoring. \




```{r,include=F}
glm.model_4_bodypart <- glm(shot.outcome.name ~ shot.body_part.name, data = df_model_4, family = binomial(link = "logit"))
summary(glm.model_4_bodypart)

#The probability of marking the head is lower than for other parts of the body.

```
```{r,include=F}
model_constant= glm(shot.outcome.name ~ 1, data = df_model_4, family = binomial(link = "logit"))
anova(model_constant,glm.model_4_bodypart,test="Chisq")
```



Testing the model with only the shot.body_part.name variable gives us a $p_{value}$ of :
\( `r round(anova(model_constant, glm.model_4_bodypart, test="Chisq")$Pr[2], digits = 3)` \). \
We reject $H_0$, the technique variable is significant.



We now want to test the model with only the shot.technique.name variable.


```{r,echo=F}

glm.model_4_technique <- glm(shot.outcome.name ~ shot.technique.name, data = df_model_4, family = binomial(link = "logit"))
summary(glm.model_4_technique)
```

The reference is backheel : all the other techniques are better, we have a lot of values close to 1, we could do a constant sub-model to see if this variable is significant. \

```{r,include=F}
model_constant= glm(shot.outcome.name ~ 1, data = df_model_4, family = binomial(link = "logit"))
anova(model_constant,glm.model_4_technique,test="Chisq")
```

We find a $p_{value}$ of \( `r round(anova(model_constant, glm.model_4_technique, test="Chisq")$Pr[2], digits = 3)` \).
We reject $H_0$, the technique variable is significant.



We are now testing the model with only the shot.type.name variable. We also run a sub-model test. \


```{r,include=F}

glm.model_4_type <- glm(shot.outcome.name ~ shot.type.name, data = df_model_4, family = binomial(link = "logit"))
summary(glm.model_4_type)

```

```{r,include=F}
anova(model_constant,glm.model_4_type,test="Chisq")
```
We find a $p_{value}$ of \( `r round(anova(model_constant, glm.model_4_type, test="Chisq")$Pr[2], digits = 3)` \).

The variable shot.type.name is significant, we reject $H_0$.



We do the same with the variable location.x : 

```{r,include=F}
glm.model_4_loc_x <- glm(shot.outcome.name ~ location.x, data = df_model_4, family = binomial(link = "logit"))
summary(glm.model_4_loc_x)
```

```{r,include=F}
anova(model_constant,glm.model_4_loc_x,test="Chisq")
```
We see a $p_{value}$ of : \( `r round(anova(model_constant, glm.model_4_loc_x, test="Chisq")$Pr[2], digits = 3)` \). <0.05 so location.x is highly significant.\


We check if the variable location.y is significant as well. \

```{r,include=F}
glm.model_4_loc_y <- glm(shot.outcome.name ~ location.y, data = df_model_4, family = binomial(link = "logit"))
summary(glm.model_4_loc_y)
```

```{r,include=F}
anova(model_constant,glm.model_4_loc_y,test="Chisq")
```

The $p_{value}$ is : \( `r round(anova(model_constant, glm.model_4_loc_y, test="Chisq")$Pr[2], digits = 3)` \). > 0.05 so location.y is not significant.\


### Testing the complete model


The model is now tested with all the following variables: shot.body_part.name,shot.technique.name,shot.type.name,location.x,location.y,under_pressure.\


```{r,include=F}
#We test now with all the variable in the model 3+under_pressure 
glm.model_5 <- glm(shot.outcome.name ~ ., data = df_model_4, family = binomial(link = "logit"))
summary(glm.model_5)
AIC_model5=AIC(glm.model_5)
R2_model5=1-(glm.model_5$deviance/glm.model_5$null.deviance)
print(R2_model5)
```

We have an $R^2$ of \( `r R2_model5` \) which is good, but it's normal because it's a model with many variables. \

We also note a low AIC, which is equal to \texttt{`r AIC_model5`}.\

## Model 5 : adding the position of the goalkeeper


We create the same model as above, but adding the position of the goalkeeper. \

### Does the position of the goalkeeper in x and y improve our results ? 

First we test the model with only the location.x.GK variable.\

```{r,echo=F}
#Create a new df with all previous variables + goalkeeper position
df_model_6 <- dplyr::select(shots,shot.outcome.name,shot.body_part.name,shot.technique.name,shot.type.name,location.x,location.y,under_pressure,location.x.GK,location.y.GK)
df_model_6$shot.outcome.name <- ifelse(df_model_6$shot.outcome.name == "Goal", 1, 0)

#Replace NA with “false” in the under_pressure column
df_model_6$under_pressure <- ifelse(is.na(df_model_6$under_pressure), FALSE, df_model_6$under_pressure)

#Note that for penalty kicks there is no value for the position of the goalkeeper on the pitch, so we set his position on his line, i.e. at x=120, in the location.x.GK column.
df_model_6$location.x.GK <- ifelse(is.na(df_model_6$location.x.GK), 125.0, df_model_6$location.x.GK)

#Testing the model with only the location.x.GK variable
glm.model_6_location.x.GK <- glm(shot.outcome.name ~ location.x.GK, data = df_model_6, family = binomial(link = "logit"))
summary(glm.model_6_location.x.GK)

AIC_model6_locx_GK=AIC(glm.model_6_location.x.GK)

print(1-(glm.model_6_location.x.GK$deviance/glm.model_6_location.x.GK$null.deviance))
```

Significant effect of goal position in x because both $p_{values}$ are lower than 0.5. \
The AIC value is low, equals to \texttt{`r AIC_model6_locx_GK`}.\

Then we do the same but with the location.y.GK variable.\

```{r,include=F}
# As before we note that for penalty kicks there is no value for the position of the goalkeeper on the pitch, so we set his position on his line, i.e. at y=40, in the location.y.GK column.
df_model_6$location.y.GK <- ifelse(is.na(df_model_6$location.y.GK), 40.0, df_model_6$location.y.GK)

#We test the model with only the variable location.y.GK
glm.model_6_location.y.GK <- glm(shot.outcome.name ~ location.y.GK, data = df_model_6, family = binomial(link = "logit"))
summary(glm.model_6_location.y.GK)

AIC_model6_locy_GK=AIC(glm.model_6_location.y.GK)


R2_model6loc_y_GK=1-(glm.model_6_location.y.GK$deviance/glm.model_6_location.y.GK$null.deviance)
print(R2_model6loc_y_GK)
```

We find that the variable for keeper position in y is significant as well.
AIC is slightly higher than for position in x, it's equal to \texttt{`r AIC_model6_locy_GK`}.\


### Complete model 



```{r,include=F}
#Testing the model with all df6 variables without interaction
glm.model_6 <- glm(shot.outcome.name ~ ., data = df_model_6, family = binomial(link = "logit"))
summary(glm.model_6)

AIC_model6=AIC(glm.model_6)

R2_model6=1-(glm.model_6$deviance/glm.model_6$null.deviance)
print(R2_model6)

```

For the model with all the preceding variables and without interaction, we find a very low AIC=\texttt{`r AIC_model6`}. \
We can conclude that this model is really good.\


## Keeping all significant variables and removing location.y

```{r,include=F}
#We create a new df without location.y
df_model_7 <- dplyr::select(shots,shot.outcome.name,shot.body_part.name,shot.technique.name,shot.type.name,location.x,under_pressure,location.x.GK,location.y.GK)
df_model_7$shot.outcome.name <- ifelse(df_model_7$shot.outcome.name == "Goal", 1, 0)

#Replace NA with “false” in the under_pressure column
df_model_7$under_pressure <- ifelse(is.na(df_model_7$under_pressure), FALSE, df_model_7$under_pressure)

#Note that for penalty kicks there is no value for the position of the goalkeeper on the pitch, so we set his position on his line, i.e. at x=120 and at y=40, in the location.x.GK column.
df_model_7$location.x.GK <- ifelse(is.na(df_model_7$location.x.GK), 120.0, df_model_7$location.x.GK)
df_model_7$location.x.GK <- ifelse(is.na(df_model_7$location.y.GK), 40.0, df_model_7$location.y.GK)

#Test the model with all variables
glm.model_7 <- glm(shot.outcome.name ~ ., data = df_model_7, family = binomial(link = "logit"))
summary(glm.model_7)

AIC_model7=AIC(glm.model_7)

R2_model7=1-(glm.model_7$deviance/glm.model_7$null.deviance)
print(R2_model7)

```

Since we found that location.y is not significant, we can remove it from the model.\ 

Without this variable, the AIC is even lower, at \texttt{`r AIC_model7`}.\

We can conclude that we have found our best model for now and it's composed of the variables : 

body_part, shot.technique, shot.type, location.x, under_pressure, location.x.GK, location.y.GK.\

## Replacing location.y

```{r, include=F}
df_model_8 <- dplyr::select(shots,shot.outcome.name,location.x,location.y,)
df_model_8 <- df_model_8 %>% mutate(distance_to_center = case_when(location.y >40 ~ location.y-40,
                                                                   location.y<=40 ~ -(location.y-40)))

df_model_8$shot.outcome.name <- ifelse(df_model_8$shot.outcome.name == "Goal", 1, 0)

glm.model_8 <- glm(shot.outcome.name ~ location.x + distance_to_center, data = df_model_8, family = binomial(link = "logit"))
summary(glm.model_8)

AIC_model8=AIC(glm.model_8)

R2_model8=1-(glm.model_8$deviance/glm.model_8$null.deviance)
print(R2_model8)
```

Seeing as location.y is a very insignificant variable, we now view offense as symmetrical with respect to the axis passing by the center point of the pitch and the penalty spots. We will now refer to location.y as the distance to center, and the new variable will vary from 0 to 40 meters. This hopefully will give more sense to the parameter, as a high distance to center means a player is very off-center.

```{r, echo = F,eval = TRUE, warning=FALSE}
part = 256

new_df_predicted <- data.frame(x = numeric(part^2),
                         y = numeric(part^2))

for (i in (1:part)){
    for (j in (1:part)){
      loc_x = ((i-1)*120 + i*120)/(2*part)
      loc_y = ((j-1)*80 + j*80)/(2*part)
      new_df_predicted[(i-1)*part+j,]$x = loc_x
      new_df_predicted[(i-1)*part+j,]$y = loc_y
  }
}

new_df_predicted <- new_df_predicted %>% mutate(location.x = x,
                                                location.y = y,
                                    distance_to_center = case_when(y >40 ~ y-40,
                                                                   y<=40 ~ -(y-40)))

new_df_predicted <- new_df_predicted %>% mutate(xg = predict(glm.model_8, new_df_predicted, type="response"))
suppressWarnings(suppressMessages(create_Pitch(opposing=TRUE, Heatmap = TRUE, df_heatmap = new_df_predicted, grass_colour = NA, line_colour = "white")))
```

The above heatmap shows what is to be expected : from a very simple model, only using the location as well as the distance to center, we can see that the model expects a player to have better scoring chances with point-blank shots rather than shots outside the penalty area. This is explained by the fact that our data set does not contain many goals made from outside this area. 

```{r,echo=F}

#We want to compare our xg obtained with model 7 with those of statsbomb :

#We have a size problem between the two vectors: there are missing values
#We obtain the indices of missing observations for all variables in model 7

indices_obs_manquantes <- which(!complete.cases(df_model_7))
indices_non_manquants <- setdiff(1:nrow(df_model_7), indices_obs_manquantes)

#Only rows with no missing observations are kept
statsbomb_xg <- shots$shot.statsbomb_xg[indices_non_manquants]

df_plot <- data.frame(statsbomb_xg = statsbomb_xg, fitted = glm.model_7$fitted.values)

#We compare both
ggplot(df_plot, aes(x = seq_along(statsbomb_xg))) +
  geom_point(aes(y = statsbomb_xg), color = "blue") +
  geom_point(aes(y = fitted), color = "red") +
  labs(title = "Comparison of statsbomb_xg and fitted values",
       x = "Observations",
       y = "Values") +
  theme_minimal()

ggplot(df_plot, aes(x = seq_along(statsbomb_xg))) +
  geom_point(aes(y = log(statsbomb_xg)), color = "blue") +
  geom_point(aes(y = log(fitted)), color = "red") +
  labs(title = "Comparison of statsbomb_xg and fitted values",
       x = "Observations",
       y = "Values") +
  theme_minimal()


#We plot one against the other: xg against our fitted values
ggplot(df_plot, aes(x = statsbomb_xg, y = fitted)) +
  geom_point() + 
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(title = "Comparison of statsbomb_xg and fitted values",
       x = "statsbomb_xg",
       y = "Fitted values") + 
  theme_minimal()  

#Same as transforming to log
ggplot(df_plot, aes(x = log(statsbomb_xg), y = log(fitted))) +
  geom_point() + 
  geom_abline(intercept = 0, slope = 1, color = "red") + 
  labs(title = "Comparison of statsbomb_xg and fitted values",
       x = "statsbomb_xg",
       y = "Fitted values") + 
  theme_minimal()  


```

We have a lot of values close to 0, so we do the log to make things clearer.

In log: there's a point (an observation) where we've overestimated the chance of scoring.
There are a few points at the bottom right where, on the contrary, we've underestimated the probability, but overall we've got a good prediction based on the Xg of bomb stats.


## Finding our best model with the AIC criterion

```{r,echo=F}
#Find the best model (that minimizes AIC)
bestmod=stepAIC(glm.model_6, trace=FALSE)
summary(bestmod)
df_model_6_best <- dplyr::select(shots,shot.outcome.name,shot.body_part.name,shot.technique.name,shot.type.name,location.x,location.x.GK)
#Note that for penalty kicks there is no value for the position of the goalkeeper on the pitch, so we set his position on his line, i.e. at x=119, in the location.x.GK column.
df_model_6_best$location.x.GK <- ifelse(is.na(df_model_6_best$location.x.GK), 119.0, df_model_6_best$location.x.GK)
#view(df_model_6_best)
```

We can see finally that y-positions are useless even for the goalkeeper, only x-positions are significant.

Our best model is composed of 5 variables : 
The technique of shot, the type of shot, the body part used and the positions in x for the player and the goalkeeper.\

# Analysis of the Model performance

Now that we have our model composed of these 5 variables, we asked ourselves if this type of model is better than the Statsbomb xG one, and overall we wanted to test the perfomance of our model !
So as we previsouly did in the "Descriptive Analysis" part of our report we are conducting a comparison on the ratio of goals. Before, we did it between data we were given, but now we will predict the xG of every shot with our best model and then compare these xG value to the Statsbomb one on graphics.

Figure \ref{fig:fig11} shows a visualization of these results.\


```{r,echo=F}
Shots_exp_predi <- subset(shots, select = c(team.name, player.name,shot.statsbomb_xg,shot.outcome.name,shot.body_part.name,shot.technique.name,shot.type.name,location.x,location.x.GK))
Shots_exp_predi <- Shots_exp_predi  %>%  mutate(player = player.name,
                                      .keep="unused")
#Note that for penalty kicks there is no value for the position of the goalkeeper on the pitch, so we set his position on his line, i.e. at x=120, in the location.x.GK column.
Shots_exp_predi$location.x.GK <- ifelse(is.na(Shots_exp_predi$location.x.GK), 120.0, Shots_exp_predi$location.x.GK)

# Calculer les probabilités prédites de réussite de chaque passe (xG)
xG_predi <- predict(bestmod, newdata = Shots_exp_predi, type = "response")

Shots_exp_predi$xG_predi <- xG_predi
Shots_exp_predi$xGdiff <- Shots_exp_predi$shot.statsbomb_xg - Shots_exp_predi$xG_predi #We create a indicator to see the difference between the StatsBomb xG and the predicted one
#We don'ttake the absolute value of this difference to keep the information of which model predict a greater probability than the others ("+" is the StatsBomb one /// "-" is our's)


#view(xG_predi)
#view(Shots_exp_predi)


df_shot_xG_succeed <- aggregate(x = Shots_exp_predi$shot.statsbomb_xg, by = list(Shots_exp_predi$player), FUN = sum)
df_shot_xGpredi_succeed <- aggregate(x = Shots_exp_predi$xG_predi, by = list(Shots_exp_predi$player), FUN = sum)

colnames(df_shot_xG_succeed) <- c("player", "xG")
colnames(df_shot_xGpredi_succeed) <- c("player", "xG_predi")
#view(df_shot_xG_succeed)
#view(df_shot_xGpredi_succeed)


df2 <- merge(df, df_shots_exp_succeed, by=c("player"))
#view(df2)
df2 <- df2 %>% mutate(xG_ratio = df_shots_exp_succeed$xG/total_shots)
#,                  .keep = "unused")

df2 <- subset(df2, select = -xG)
df3 <- merge(df2, df_shot_xG_succeed, by=c("player"))
#view(df3)
df3 <- merge(df3, df_shot_xGpredi_succeed, by=c("player"))


df3 <- df3 %>% mutate(xG_ratio = df_shot_xG_succeed$xG/total_shots,
                      xG_predi_ratio = df_shot_xGpredi_succeed$xG_predi/total_shots,
                        total_shots=total_shots,
                        .keep = "unused")

print(length(df3$player))

df3 <- df3[df3$total_shots>=1,] # Removing players that attempted less than 1 shots
#view(df3)
print(length(df3$player)) # 398 players attempted at least 1 shot in the WC

#We now have a df containing every player that shots at least once in the World cup, associated with their xg
```

```{r,echo=F}
pays = c("France", "England", "Spain", "Switzerland") #Select teams we wish to observe
pays <- melt(df3$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #Select all teams
#view(df3)
df3_player <- df3[df3$team %in% pays,]
#view(df3_player)

df3_player$groupv2 <- ifelse(df3_player$team %in% pays_gs, "8-GS",
                              ifelse(df3_player$team %in% pays_r16, "7-R16",
                                     ifelse(df3_player$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(df3_player$team == "Australia", "4th",
                                                   ifelse(df3_player$team == "Sweden", "3rd",
                                                          ifelse(df3_player$team == "England", "2nd",
                                                                 ifelse(df3_player$team=="Spain","1st","NA")))))))


mean_shots_nb = mean(df3_player$total_shots)
mean_goals_nb = mean(df3_player$total_goals)
mean_shots_completion_ratio = mean_goals_nb/mean_shots_nb
mean_shots_xG = mean(df3_player$xG)
mean_shots_xG_ratio = mean_shots_xG/mean_shots_nb
mean_shots_xG_predi = mean(df3_player$xG_predi)
mean_shots_xG_predi_ratio = mean_shots_xG_predi/mean_shots_nb

df3_player[nrow(df3_player) + 1,] <- list("Mean Player", "Mean Player", mean_shots_nb, mean_shots_completion_ratio, mean_goals_nb, mean_shots_xG_ratio, mean_shots_xG, mean_shots_xG_predi, mean_shots_xG_predi_ratio, "Mean Player")


#view(df3_player)
```


```{r fig11 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fi11}StatsBomb xG ratio on our xG ratio based on bestmod for players",fig.height=4.5,fig.align='center'}
p4 <-ggplot(df3_player, aes(xG_predi_ratio, xG_ratio, size = total_shots, label = player, colour=groupv2)) + geom_point(shape=21) + geom_abline(intercept = 0, slope = 1, linetype = "dashed") + scale_color_viridis(discrete=TRUE) + geom_point(data = df3_player[df3_player$team == "Mean Player", ], aes(xG_predi_ratio, xG_ratio, size = total_shots), shape = 4, color = "red4", stroke = 1) + labs(x = "Our Model's Prediction of the xG's ratio", y = "StatsBomb's Model xG's Ratio", fill = "World Cup's Stages", color = "World Cup's Stages")

# Display “Mean Player” with a red cross
#p4 <- p4 + geom_point(data = df3_player[df3_player$team == "Mean Player", ], aes(xG_predi_ratio, xG_ratio,  size = total_shots), shape = 4, color = "red4", stroke = 1)

ggplotly(p4)
p4
```

We notice that if we display the ratio of the xG we predicted to those from Statsbomb, the results vary very little, which means our model is quite close to that of Statsbomb! However, to accurately compare our model with Statsbomb's, we should display the xG for each shot. 

This is what the \ref{fig:fig12} below shows:\

```{r fig12 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig12}StatsBomb xG  on our xG  based on bestmod for every shots in the World Cup",fig.height=4.5,fig.align='center'}
pp <-ggplot(Shots_exp_predi, aes(xG_predi, shot.statsbomb_xg, color=xGdiff)) + geom_point(shape=21) + geom_abline(intercept = 0, slope = 1, linetype = "dashed") +   scale_color_gradientn(colors = c("darkblue", "blue3", "blue2", "darkolivegreen3", "gold" ,"darkgoldenrod2", "darkred")) + labs(x = "Our Model's Prediction of the xG's ratio", y = "StatsBomb's Model xG's Ratio", fill = "Differences of xG", color = "Differences of xG")

ggplotly(pp)
pp
```
We can see that for the vast majority of shots, our models are similar! However, we do notice some outliers, which can be explained by displaying, for example, the type of shot, whether it is a free kick, an open play shot, a penalty, a corner, etc. This is shown in \ref{fig:fig13} below:\

```{r,echo=F}
#view(Shots_exp_predi)

df_xGdiff <- Shots_exp_predi[order(Shots_exp_predi$xGdiff),] #We sort the dataframe in ascending order of the difference between statsbomb's xG and our xG
#view(df_xGdiff)

#We retrieve the first x rows (the largest values)
x <- 50
df_top_xGdiff <- head(df_xGdiff, x)
df_low_xGdiff <- tail(df_xGdiff, x)


#view(df_top_xGdiff)
#view(df_low_xGdiff)

df_top_xGdiff$source <- "StatsBomb xG > Our's xG"
df_low_xGdiff$source <- "StatsBomb xG < Our's xG"

# Combiner les deux DataFrames
df_combined <- rbind(df_top_xGdiff, df_low_xGdiff)
```


```{r fig13 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig13}StatsBomb xG  on our xG  based on bestmod and the type of shots for the 100 shots with the higher xG difference",fig.width=8,fig.asp=0.7,fig.align='center'}
#To understand where can these differences came from we plot the shots with the highest differences and we print the type of shots
p_combined <- ggplot(df_combined, aes(x = xG_predi, y = shot.statsbomb_xg, label = shot.type.name, color = source)) +
  geom_text(aes(fontface=3)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") + labs(x = "Our Model's Prediction of the xG's ratio", y = "StatsBomb's Model xG's Ratio", color = "Difference xGs")

ggplotly(p_combined)
p_combined
```


```{r, eval=F, echo=F}
#To understand where can these differences came from we plot the shots with the highest differences and we print the body part with which the shot was taken
p_combined2 <- ggplot(df_combined, aes(x = xG_predi, y = shot.statsbomb_xg, label = shot.body_part.name, color = source)) +
  geom_text(aes(fontface=3)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  labs(color = "Source") + 
  theme_minimal()

ggplotly(p_combined2)

#To understand where can these differences came from we plot the shots with the highest differences and we print the technique of the shots


p_combined3 <- ggplot(df_combined, aes(x = xG_predi, y = shot.statsbomb_xg, label = shot.technique.name, color = source)) +
  geom_text(aes(fontface=3)) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  labs(color = "Source") +
  theme_minimal()

ggplotly(p_combined3)
```

We can see a clear example of bias in our model: the 'Corner' at the bottom right of the graph, where our model predicts an xG of 1 and Statsbomb predicts an xG of almost 0. In fact, during the competition, there was an in-swinging corner that was attempted and succeeded, so our model, having been trained only on this competition, now classifies all in-swinging corners as shots with a goal probability of more than 0.99. On the other hand, Statsbomb's model, likely trained on many more competitions, has enough historical data to see that this rare type of shot very seldom results in a goal, thus it assigns a probability close to 0.

For the other points, it is more complicated to explain. Indeed, they are almost all open play shots, so many other parameters come into play. Our model only considers five: the type of shot (penalty, free kick, open play, etc.), the shot technique (normal, lob, volley, bicycle kick, etc.), the part of the body used for the shot, and the distance to the goal of the player and the goalkeeper. Therefore, we are probably missing some information, such as the shot angle or whether the player was under pressure, or we need more data from women's football tournaments to refine our model and smooth out the peculiarities of this competition and the impact of anomalies, like the in-swinging corner, by analyzing more than 64 matches.


\clearpage

# Another Model : Expected Pass

Having analyzed the Expected Goal (xG) values provided by StatsBomb and developed our own xG prediction model, we questioned whether StatsBomb offered a similar indicator for passes—an Expected Pass (xPass) metric, which predicts the probability of a pass being successful based on various parameters. Surprisingly, they do not have such an indicator. Consequently, we decided to implement one using the same methodology employed for our xG model. We chose to create this model for passes because a football team's performance is strongly correlated with goals, as goals ultimately determine the outcome of matches. However, when observing football matches, it is evident that some teams struggle to create scoring opportunities and goal situations due to their inability to execute passes between the lines, through passes, player combinations, technical skills, or dribbles. We have noted that teams reaching the quarterfinals generally exhibited better xG and True Goal Ratios compared to others, with some exceptions like Colombia. This model allows us to analyze another dimension of the complex sport of football.

For instance, did Colombia excel in the area of play construction compared to their opponents, thereby enabling them to reach the quarterfinals despite having similar xG and True Goal Ratios? Moreover, do we observe the same trend regarding goal-scoring capabilities in the top teams? These questions prompted us to develop this model for analyzing passes.

## Quick Descriptive Analysis

```{r, eval=T, echo=F}
Passes_All <- Events %>% filter(type.name=="Pass")
#view(Passes_All)
Passes_All <- Passes_All  %>%  mutate(team = gsub(" ","",team.name),
                                      .keep="unused")

#view(Passes_All)
Pass <- subset(Passes_All, select = c(pass.outcome.name,team, player.name))
```

```{r, eval=T, echo=F}
Pass <- Pass  %>%  mutate(player = player.name,
                                      outcome = case_when(is.na(pass.outcome.name) ~ 1,
                                                         !is.na(pass.outcome.name) ~ 0),
                                      team = gsub("Women's","",team),
                                      .keep="unused")



#View(Passes_All)
```

### Player-by-player visualization

```{r, eval=T, echo=F, include=FALSE} 
df_pass_total <- aggregate(x = Pass$outcome, by = list(Pass$player), FUN = length)
df_pass_succeed <- aggregate(x = Pass$outcome, by = list(Pass$player), FUN = sum)

df <- merge(df_pass_succeed, df_pass_total, by=c("Group.1")) %>% mutate(player = Group.1,
                                                                               completed_pass = x.x,
                                                                               total_pass = x.y,
                                                                               .keep="unused")

df1 <- subset(Pass, selec = c(player, team))
#view(df1)

df <- merge(df1, df, by=c("player"))
#view(df)
df <- df %>% distinct()

df <- df %>% mutate(pass_completion_ratio = completed_pass/total_pass,
                    total_pass = total_pass,
                    total_pass_completed = completed_pass,
                    .keep = "unused")

print(length(df$player))  # 612 players attempted at least 1 pass in the WC
#view(df)

df <- df[df$total_pass>=4,] # Removing players that attempted less than 4 passes

print(length(df$player)) # 588 players attempted at least 4 pass in the WC

#We now have a df containing every player that shot at least once in the World cup, associated with how many shots they attempted, and their ratio of successful shots

#pays = c("France", "England", "Spain", "Switzerland") #Select teams we wish to observe
pays <- melt(df$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #Select all teams

df_player <- df[df$team %in% pays,]


df_player$groupv2 <- ifelse(df_player$team %in% pays_gs, "8-GS",
                              ifelse(df_player$team %in% pays_r16, "7-R16",
                                     ifelse(df_player$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(df_player$team == "Australia", "4th",
                                                   ifelse(df_player$team == "Sweden", "3rd",
                                                          ifelse(df_player$team == "England", "2nd",
                                                                 ifelse(df_player$team=="Spain","1st","NA")))))))

mean_pass_completion_ratio = mean(df_player$pass_completion_ratio)
mean_pass_nb = mean(df_player$total_pass)
mean_pass_completed_nb = mean(df_player$total_pass_completed)

df_player[nrow(df_player) + 1,] <- list("Mean Player", "Mean Player", mean_pass_nb, mean_pass_completion_ratio, mean_pass_completed_nb, "Mean Player")
#view(df_player)
```


```{r fig14 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig14}True Pass ratio on the Total Number of Pass for every player with 5 pass or more during the WC",fig.width=10,fig.asp=0.8,fig.align='center'}
p6 <- ggplot(df_player, aes(total_pass, pass_completion_ratio, label = paste("player =", player, ", team =", team ), color = groupv2)) + geom_point() +  scale_color_viridis(discrete=TRUE)

p6 <- p6 + geom_point(data = df_player[df_player$team == "Mean Player", ], aes(total_pass, pass_completion_ratio),  size = 2, shape = 4, color = "red4", stroke = 1) + labs(x = "Total of Pass during the World Cup", y = "True Successful Pass Ratio for Players", color = "World Cup Stages") 

ggplotly(p6)
p6
```

This graph illustrates the ratio of successful passes to the total number of passes. Once again, a striking difference is observed: players from countries that reached the quarterfinals are positioned at the top of the graph, with a successful pass ratio exceeding 60%, whereas teams eliminated in the round of 16 and group stages exhibit considerably lower pass ratios. It is noteworthy that within these less successful teams, there are still a significant number of players with pass ratios as high as those in the top-performing teams. This suggests potential disparities within these countries, likely indicating a higher level of heterogeneity in player performance.\


### Team-by-team visualization

```{r, eval=T, echo=F, include=FALSE} 
df_pass_total <- aggregate(x = Pass$outcome, by = list(Pass$player), FUN = length)
df_pass_succeed <- aggregate(x = Pass$outcome, by = list(Pass$player), FUN = sum)

df <- merge(df_pass_succeed, df_pass_total, by=c("Group.1")) %>% mutate(player = Group.1,
                                                                               completed_pass = x.x,
                                                                               total_pass = x.y,
                                                                               .keep="unused")

df1 <- subset(Pass, selec = c(player, team))
#view(df1)

df <- merge(df1, df, by=c("player"))
#view(df)
df <- df %>% distinct()

df <- df %>% mutate(pass_completion_ratio = completed_pass/total_pass,
                    total_pass = total_pass,
                    total_pass_succeed = completed_pass,
                    .keep = "unused")

print(length(df$player))
#view(df)

df <- df[df$total_pass>=1,] # Removing players that attempted less than 1 pass

print(length(df$player)) # 612 players attempted at least 1 pass in the WC

#We now have a df containing every player that shots at least once in the World cup, associated with how many shots they attempted, and their ratio of successful shots

#pays = c("France", "England", "Spain", "Switzerland") #Countries we wish to observe
#pays = c("Switzerland")
pays <- melt(df$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #select all teams

#view(pays)

df_pays <- df[df$team %in% pays,]

for (i in 0:length(pays))
  df_pays[nrow(df_pays) + 1,] <- list(paste("",pays[i]), paste("", pays[i]), 
                                      sum(df_pays$total_pass[df_pays$team == pays[i]]), 
                                      0,
                                      sum(df_pays$total_pass_succeed[df_pays$team == pays[i]]))

df_pays <- subset(df_pays, select = -c(player))

n <- length(pays)

#view(df_pays)

dftail <- df_pays[(nrow(df_pays)-(n - 1)):nrow(df_pays),] %>% mutate(team = gsub(" ","",team),
                                                                     pass_completion_ratio = total_pass_succeed/total_pass,
                                                                     total_pass = total_pass,
                                                                     total_pass_succeed = total_pass_succeed,
                                                                     .keep = "unused")

#pays_gs <- pays_gs %>% mutate(team=gsub(" ","",x))

dftail$groupv2 <- ifelse(dftail$team %in% pays_gs, "8-GS",
                              ifelse(dftail$team %in% pays_r16, "7-R16",
                                     ifelse(dftail$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(dftail$team == "Australia", "4th",
                                                   ifelse(dftail$team == "Sweden", "3rd",
                                                          ifelse(dftail$team == "England", "2nd",
                                                                 ifelse(dftail$team=="Spain","1st","NA")))))))


mean_pass_completion_ratio_pays = mean(dftail$pass_completion_ratio)
mean_pass_nb_pays = mean(dftail$total_pass)
mean_pass_completed_nb_pays = mean(dftail$total_pass_succeed)

dftail[nrow(dftail) + 1,] <- list("Mean Country", mean_pass_nb_pays, mean_pass_completion_ratio_pays, mean_pass_completed_nb_pays, "Mean Country")

#view(dftail)
```


```{r fig15 ,echo=F,eval=TRUE,fig.cap="\\label{fig:fig15}True Pass ratio on the Total Number of Pass for every teams during the WC",fig.width=10,fig.asp=0.8,fig.align='center'}
p7 <- ggplot(dftail, aes(total_pass, pass_completion_ratio, label=team, color = groupv2)) + geom_text(aes(fontface=2)) + scale_color_viridis(discrete=TRUE)

p7 <- p7 + geom_point(data = dftail[dftail$team == "Mean Country", ], aes(total_pass, pass_completion_ratio),  size = 2, shape = 4, color = "red4", stroke = 1) + labs(x = "Total of Pass during the World Cup", y = "True Successful Pass Ratio for Teams", color = "World Cup Stages") 

ggplotly(p7)
p7
```

We examined the statistics by country and observed a consistent trend with a notable distinction for Spain, which made significantly more passes than other teams that played a similar number of matches, such as Sweden, Australia, and England, while also achieving a higher pass ratio. Combining this with shooting data, where Spain had many more shots than its competitors, it reaffirms that Spain controlled their matches more effectively, with higher possession and more opportunities. This observation likely extends to England, the other finalist team. For Sweden and Australia, the third and fourth place teams respectively, we can infer that their passes were more risky, potentially involving more line-breaking passes or through balls. Colombia, once again, is positioned among teams that were eliminated earlier and aligns with the average performance level of the competition.

To provide deeper interpretation and validate our analyses with these visualizations, we will now focus on the Expected Pass (xPass) indicator, particularly the difficulty of the passes.\

## Implementing the Pass model we created

```{r, eval=T, echo=F, include=FALSE} 
Passes_All <- free_allevents(AllMatches) %>% filter(type.name=="Pass")
Passes_Cleaned <- cleanlocations(Passes_All)
#View(Passes_Cleaned)
```

```{r, eval=T, echo=F, include=FALSE}
# Let's try to predict whether a pass will be completed or not.
# We will be looking at several different parameters for this model : the length, the angle, and whether a pass is under pressure or not. We also added the seconds, to verify the relevance of the results.

dfmod <- data.frame(Passes_Cleaned['pass.length'],Passes_Cleaned['pass.angle'],Passes_Cleaned['pass.outcome.name'], Passes_Cleaned['under_pressure'], Passes_Cleaned['second'], Passes_Cleaned['location.x'], Passes_Cleaned['location.y'])

#Let's transform the data using the mutate() and case_when() functions. Also, we are renaming some of the attributes to clearer names.

dfmod <- dfmod  %>%  mutate(length = pass.length,
                      angle = pass.angle,
                      outcome = case_when(is.na(pass.outcome.name) ~ 1,
                                         !is.na(pass.outcome.name) ~ 0),
                      under_pressure = case_when(is.na(under_pressure) ~ 0,
                                                !is.na(under_pressure) ~ 1))

#Transformer la variable angle

# Removing unnecessary columns

dfmod <- subset(dfmod, select = -c(pass.length, pass.angle, pass.outcome.name))
#view(dfmod)

# Setting up the model

m1 <- glm(outcome ~ length*angle*under_pressure*second + location.x + location.y, data=dfmod)

# Removing unnecessary parameters from the model

mBIC <- step(m1, backward=TRUE, k=log(nrow(dfmod)), trace=FALSE)
summary(mBIC)

# The BIC step procedure removed the "second" (this one is no surprise...) and "angle" parameters from the model.
```


```{r, eval=T, echo=F, include=FALSE}
R2 = 1 - 465.61/507.53
print(R2)
# R2 = 0.0825961, not very good by any means (with passes from France only, R2 = 0.07331737 with every pass from WC2023F)
```

```{r, eval=T, echo=F, include=FALSE}
#The location.y parameter was not considered relevant enough for the model. Adding the distance to center, as the location.y attribute should have a symmetrical impact along the halfline (the ones passing through the goalposts) from a strategical standpoint. 

dfmod <- dfmod %>% mutate(distance_to_center = case_when(location.y >40 ~ location.y-40,
                                                   location.y<=40 ~ -(location.y-40)))

dfmod <- subset(dfmod, select = -c(second, location.y))

#view(dfmod)
```



```{r, eval=T, echo=F, include=FALSE}
# There seems to be some non-linear dependencies within the model, as our barplot suggested.

dfmod <- dfmod %>% mutate(length_sq = length**2)

mtest <- glm(outcome ~ location.x*distance_to_center*length + angle * location.x + angle * distance_to_center + length*angle*under_pressure*length_sq - length:length_sq - length:under_pressure:length_sq, data=dfmod)
summary(mtest)

mfinal <- step(mtest, backward=TRUE, k=log(nrow(dfmod)),trace=FALSE)
summary(mfinal)

# Piste de Sébastien : Modèle de xP, xG, comparer résultat de regression logistique à un modèle de machine learning (SVM, Random forest, ...)
```

```{r, eval=T, echo=F, include=FALSE}
R2 <- 1 - 421.78/507.53
print(R2)
```

```{r, eval=T, echo=F, include=FALSE}
#We predict the results for each pass
xP <- predict(mfinal, newdata = dfmod, type = "response")

# We add the xP for all passes 
dfmod$xP <- xP

print(dfmod)
```

```{r, eval=T, echo=F, include=FALSE}
Pass_exp <- subset(Passes_All, select = c(team.name, player.name))
#view(Pass_exp)

Pass_exp <- Pass_exp  %>%  mutate(player = player.name,
                                      team = gsub(" Women's","",team.name),
                                      .keep="unused")

# Calculate the predicted probability of success for each pass (xP)
Pass_exp$xP <- xP
Pass_exp$xPdiff <- Pass_exp$xP - Pass_exp$xP
#view(Pass_exp)

df_pass_exp_succeed <- aggregate(x = Pass_exp$xP, by = list(Pass_exp$player), FUN = sum)
#df_pass_exp_succeed <- df_pass_exp_succeed %>% mutate(player=Group.1, 
#                               df_pass_exp_ratio=df_pass_exp_succeed/df_pass_exp_total)

#view(df_pass_exp_succeed)
colnames(df_pass_exp_succeed) <- c("player", "xP")
#view(df)
#view(df_pass_exp_succeed)
df2p <- merge(df, df_pass_exp_succeed, by=c("player"))

df2p <- df2p %>% mutate(xP_ratio = df_pass_exp_succeed$xP/total_pass,
                        total_pass=total_pass,
                        .keep = "unused")
#view(df2p)

print(length(df2p$player))

df2p <- df2p[df2p$total_pass>=1,] # Removing players that attempted less than 1 pass
#view(df2p)
print(length(df2p$player)) # 612 players attempted at least 1 pass in the WC

#We now have a df containing every player that shots at least once in the World cup, associated with their xP
```
### Comparaison xP to True Pass Ratio visualization

#### For hard shots

```{r, eval=T, echo=F, include=FALSE}
#We want to keep only the 25% of passes with the lowest xP (i.e. the hardest).
q1<-quantile(df2p$xP_ratio,0.25)
print(q1)
df2p_hard <- df2p[df2p$xP_ratio<=q1,] # Keeping only the 25% hardest passes.
#view(df2p_hard)
```
```{r, eval=T, echo=F}
#pays = c("France", "England", "Spain", "Switzerland") #Select teams we wish to observe
pays <- melt(df2p_hard$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #Select all teams
#view(df2p_player)
#view(df2p)

df2p_player <- df2p_hard[df2p_hard$team %in% pays,]
#view(df2p_player)

df2p_player$groupv2 <- ifelse(df2p_player$team %in% pays_gs, "8-GS",
                              ifelse(df2p_player$team %in% pays_r16, "7-R16",
                                     ifelse(df2p_player$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(df2p_player$team == "Australia", "4th",
                                                   ifelse(df2p_player$team == "Sweden", "3rd",
                                                          ifelse(df2p_player$team == "England", "2nd",
                                                                 ifelse(df2p_player$team=="Spain","1st","NA")))))))

mean_pass_completion_ratio = mean(df2p_player$pass_completion_ratio)
mean_pass_nb = mean(df2p_player$total_pass)
mean_pass_success=mean(df2p_player$total_pass_succeed)
mean_pass_xP = mean(df2p_player$xP)
mean_pass_xP_ratio = mean(df2p_player$xP/df2p_player$total_pass)

df2p_player[nrow(df2p_player) + 1,] <- list("Mean Player", "Mean Player", mean_pass_nb, mean_pass_completion_ratio, mean_pass_success, mean_pass_xP, mean_pass_xP_ratio)

#view(df2p_player)

p8 <- ggplot(df2p_player, aes(xP_ratio, pass_completion_ratio, size = total_pass, label = player, colour=groupv2)) + 
    geom_point(shape=21) + scale_color_viridis(discrete=TRUE) + geom_abline(intercept = 0, slope = 1, linetype = "dashed")

p8 <- p8 + geom_point(data = df2p_player[df2p_player$team == "Mean Player", ], aes(xP_ratio, pass_completion_ratio,  size = total_pass), shape = 4, color = "red4", stroke = 1)

ggplotly(p8)
```   


#### For easy shots

```{r, eval=T, echo=F, include=FALSE}
#We want to keep only the 25% of passes with the highest xP (i.e. the easiest).
q3<-quantile(df2p$xP_ratio,0.75)
print(q3)
df2p_easy <- df2p[df2p$xP_ratio>=q3,] # Keeping only the 25% easiest passes.
#view(df2p_easy)
```

```{r, eval=T, echo=F, include=FALSE}
#pays = c("France", "England", "Spain", "Switzerland") #Select teams we wish to observe
pays <- melt(df2p_easy$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #Select all teams
#view(df2p_player)
#view(df2p)

df2p_player <- df2p_easy[df2p_easy$team %in% pays,]
#view(df2p_player)

df2p_player$groupv2 <- ifelse(df2p_player$team %in% pays_gs, "8-GS",
                              ifelse(df2p_player$team %in% pays_r16, "7-R16",
                                     ifelse(df2p_player$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(df2p_player$team == "Australia", "4th",
                                                   ifelse(df2p_player$team == "Sweden", "3rd",
                                                          ifelse(df2p_player$team == "England", "2nd",
                                                                 ifelse(df2p_player$team=="Spain","1st","NA")))))))

mean_pass_completion_ratio = mean(df2p_player$pass_completion_ratio)
mean_pass_nb = mean(df2p_player$total_pass)
mean_pass_success=mean(df2p_player$total_pass_succeed)
mean_pass_xP = mean(df2p_player$xP)
mean_pass_xP_ratio = mean(df2p_player$xP/df2p_player$total_pass)

df2p_player[nrow(df2p_player) + 1,] <- list("Mean Player", "Mean Player", mean_pass_nb, mean_pass_completion_ratio, mean_pass_success, mean_pass_xP, mean_pass_xP_ratio)

#view(df2p_player)

p9 <- ggplot(df2p_player, aes(xP_ratio, pass_completion_ratio, size = total_pass, label = player, colour=groupv2)) + 
    geom_point(shape=21) + scale_color_viridis(discrete=TRUE) + geom_abline(intercept = 0, slope = 1, linetype = "dashed")

p9 <- p9 + geom_point(data = df2p_player[df2p_player$team == "Mean Player", ], aes(xP_ratio, pass_completion_ratio,  size = total_pass), shape = 4, color = "red4", stroke = 1)

ggplotly(p9)
```
#### For all shots

```{r, eval=T, echo=F, include=FALSE}
#We know do the same things but for all passes not only the easiest and the hardest.

#pays = c("France", "England", "Spain", "Switzerland") #Select teams we wish to observe
pays <- melt(df2p$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #Select all teams
#view(df2p_player)
#view(df2p)

df2p_player <- df2p[df2p$team %in% pays,]
#view(df2p_player)

df2p_player$groupv2 <- ifelse(df2p_player$team %in% pays_gs, "8-GS",
                              ifelse(df2p_player$team %in% pays_r16, "7-R16",
                                     ifelse(df2p_player$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(df2p_player$team == "Australia", "4th",
                                                   ifelse(df2p_player$team == "Sweden", "3rd",
                                                          ifelse(df2p_player$team == "England", "2nd",
                                                                 ifelse(df2p_player$team=="Spain","1st","NA")))))))

mean_pass_completion_ratio = mean(df2p_player$pass_completion_ratio)
mean_pass_nb = mean(df2p_player$total_pass)
mean_pass_success=mean(df2p_player$total_pass_succeed)
mean_pass_xP = mean(df2p_player$xP)
mean_pass_xP_ratio = mean(df2p_player$xP/df2p_player$total_pass)

df2p_player[nrow(df2p_player) + 1,] <- list("Mean Player", "Mean Player", mean_pass_nb, mean_pass_completion_ratio, mean_pass_success, mean_pass_xP, mean_pass_xP_ratio)

#view(df2p_player)

p10 <- ggplot(df2p_player, aes(xP_ratio, pass_completion_ratio, size = total_pass, label = player, colour=groupv2)) + 
    geom_point(shape=21) + scale_color_viridis(discrete=TRUE) + geom_abline(intercept = 0, slope = 1, linetype = "dashed")

p10 <- p10 + geom_point(data = df2p_player[df2p_player$team == "Mean Player", ], aes(xP_ratio, pass_completion_ratio,  size = total_pass), shape = 4, color = "red4", stroke = 1)

ggplotly(p10)
```


###Comparaison xP to True Pass Ratio visualization - Pays par Pays


```{r, eval=T, echo=F, include=FALSE}
#pays2 = c("France", "England", "Spain", "Switzerland") #Countries we wish to observe
pays2 <- melt(df2p$team) %>% distinct() %>% as.list() %>% unlist() %>% sort() %>% as.list() #select all teams

#view(pays2)

df2p_pays <- df2p[df2p$team %in% pays2,]
#view(df2p_pays)

for (i in 0:length(pays2))
  df2p_pays[nrow(df2p_pays) + 1, ] <- list(paste("",pays2[i]), paste("", pays2[i]),
                                        sum(df2p_pays$total_pass[df2p_pays$team == pays2[i]]),
                                        sum(df2p_pays$pass_completion_ratio[df2p_pays$team == pays2[i]]),
                                        sum(df2p_pays$total_pass_succeed[df2p_pays$team == pays2[i]]),
                                        sum(df2p_pays$xP[df2p_pays$team == pays2[i]]),
                                        0)
  
df2p_pays <- subset(df2p_pays, select = -c(player))

#view(df2p_pays)
n <- length(pays2)

df2ptail <- df2p_pays[(nrow(df2p_pays)-(n - 1)):nrow(df2p_pays),]  %>% mutate(team = gsub(" ","",team),
                                                                    pass_completion_ratio = total_pass_succeed/total_pass,
                                                                    total_pass = total_pass,
                                                                    total_pass_succeed = total_pass_succeed,
                                                                    xP = xP,
                                                                    xP_ratio = xP/total_pass,
                                                                    .keep = "unused")
#view(df2ptail)

df2ptail$groupv2 <- ifelse(df2ptail$team %in% pays_gs, "8-GS",
                              ifelse(df2ptail$team %in% pays_r16, "7-R16",
                                     ifelse(df2ptail$team %in% pays_quarterfinals, "6-Quarterfinals",
                                            ifelse(df2ptail$team == "Australia", "4th",
                                                   ifelse(df2ptail$team == "Sweden", "3rd",
                                                          ifelse(df2ptail$team == "England", "2nd",
                                                                 ifelse(df2ptail$team=="Spain","1st","NA")))))))


mean_pass_completion_ratio_pays = mean(df2ptail$pass_completion_ratio)
mean_pass_nb_pays = mean(df2ptail$total_pass)
mean_pass_success_pays = mean(df2ptail$total_pass_succeed)
mean_pass_xP_pays = mean(df2ptail$xP)
mean_pass_xP_ratio_pays = mean(df2ptail$xP/df2ptail$total_pass)

df2ptail[nrow(df2ptail) + 1,] <- list("Mean Country", mean_pass_nb_pays, mean_pass_completion_ratio_pays, mean_pass_success_pays, mean_pass_xP_pays, mean_pass_xP_ratio_pays, "Mean Country")



#view(df2ptail)


p11 <- ggplot(df2ptail, aes(xP_ratio, pass_completion_ratio, size = total_pass, label=team, color = groupv2)) + geom_text(aes(fontface=3)) + geom_abline(intercept = 0, slope = 1, linetype = "dashed") + scale_color_viridis(discrete=TRUE)

p11 <- p11 + geom_point(data = df2ptail[df2ptail$team == "Mean Country", ], aes(xP_ratio, pass_completion_ratio,  size = total_pass), shape = 4, color = "red4", stroke = 1)

ggplotly(p11)
```